### 2019-04-19 template/articles/article_fan.tex
First Template File Created

\input{../../preambles/article_fan_pre_main.tex}

\begin{document}

\title{Paper Name}

\author{Fan Wang\thanks{Department of Economics, University of Houston (email: fwang26@uh.edu)}}

\maketitle
\begin{abstract}
\singlespacing This is my abstract \end{abstract}
\vfill
\pagebreak{}

\section{Introduction}

\section{Main}

\section{Conclusion}

\pagebreak
\input{../../cite/endcite_fan_main.tex}
\pagebreak

\appendix
\section{Data Figures and Tables}

\end{document}