### 2019-04-19 sections/data/data_one.tex
Section based compilation

\label{subsec:dataone}

The survey also captures some aspects of the pecuniary \FCFs for acquiring each loan. There was a significant reduction in these costs after 2001 for \NLC \BBBB. Most of the drop is due to the prevalence of the Million Baht Fund program. Loans from the Million Baht Fund had fees and transportation costs similar to \BBBB \LLCLY, which is likely due to its localized and village-committee based administration.

Fees and transportation costs for the BAAC also went down over time. These might have been driven down by improvements in the BAAC itself or reductions in transportation and communication costs. Table \ref{tab:FCBorrDetailType} shows that costs at the BAAC, however, are still twice as large as the average costs reported for \BBBB from relatives or moneylenders. Strikingly, the fees and transport costs on the few reported commercial bank loans are on average seven times larger than those for BAAC loans, and almost 35 times larger than the costs for Million Baht Fund loans or \LLC loans.