### 2019-04-19 sections/data/data_two.tex

\label{subsec:datatwo}

In the literature on \LLC credit markets, repayment rates for \LLC \BBBB have often been found to be high \autocite{morduch_microfinance_1999}. Recently, evidence from a number of microfinance studies shows that repayment rates for \NLC \BBBB have also been found to be high \autocite{buera_macroeconomics_2012}. Consistent with these findings, in the Townsend Thai Monthly Survey villages, both \NLC and \LLC repayments are high. These on-time repayment rates are directly calculated from the data: every month, new loans that households take out are recorded and the month in which repayment should be completed is also recorded; then every month, repayment amounts of each loan is tracked until full repayment.

Specifically, the eventual full repayment rate for \NLC loans is 97\% and for \LLC loans is 95\%. Interestingly, households that do not fully pay back their loans still pay a significant proportion of their debts, as shown in Table \ref{tab:Repay_notFull}. Taking partial repayments into consideration, I find that 98.2\% of every Baht of principal lent is paid back for \NLC loans, and 96.5\% of every Baht of principal lent is paid back for \LLC loans.