### 2019-04-19 sections/model/model_two.tex

\label{subsec:modeltwo}

At the beginning of each period, for each household, the household \PDT \(\A\), \SKE \(\SHK\), and \PYS \(\K\), which is chosen previously by the household based on expected productivity, jointly determine the \ICM of the household in the current period:\footnote{The model focuses on financial (safe asset) and \PYS  (risky asset) choices. Labor supply for the household firm is inelastic and captured by \(\A\).}
\begin{equation}
  \label{eq:ProductionFunction}
  \Y=\exp\left(\A+\SHK\right)\cdot \K^{\PALPHA}
\end{equation}