### 2019-04-19 sections_sandbox/estimate_sandbox.tex

\documentclass[12pt,english]{article}
\input{../fragments/preamble_main.tex}
\begin{document}

\input{../sections/esti/esti_main.tex}
% \input{../sections/esti/esti_one.tex}
% \input{../sections/esti/esti_two.tex}

\input{../fragments/cite/cite_end.tex}
\end{document}