### 2019-04-19 sections_blank/tabfig/main_esti.tex
Empty File

% \FIGESTIFIT{../graphs/estimation}
% \input{../preamble/group_d_data_pick/PARAMS.tex} # top of draft used by data tables as well
% \input{../preamble/group_d_data_pick/DATASIMUPROBJ.tex}  # top of draft used by data tables as well
% \FIGESTIMATES
\input{../preamble/group_d_data_pick/var/PARAMSFY1000early.tex}
% 2. Get SD
\input{../preamble/group_d_data_pick/var/PARAMSFY10early.tex}
% 3. Get SD for FB
\input{../preamble/group_d_data_pick/var/PARAMSFY500early.tex}
% 4a. sds
\input{../preamble/group_d_data_pick/var/PARAMSLV15.tex}
% 4a. sds
\input{../preamble/group_d_data_pick/var/PARAMSLV105.tex}
% 4a. mu A
\input{../preamble/group_d_data_pick/var/PARAMSLV500.tex}
% 4a. Kappa 99-01 mean + sd for all kappa
\input{../preamble/group_d_data_pick/var/PARAMSLV10.tex}
% 4c. Kappa 02-09, overrid mean 02-09
\input{../preamble/group_d_data_pick/var/PARAMSLV1000.tex}
% 4b. beta NE mean
\input{../preamble/group_d_data_pick/var/PARAMSLV100.tex}
\input{../preamble/group_d_data_pick/var/PARAMSLV50.tex}
% finalized Production function estimates o11 and real
\input{../preamble/group_d_data_pick/prod_var/PARAMS_adist_o11_rel.tex}

\FIGESTIMATESFY{Parameter Estimates}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FRACTIONS Borrowing
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Borrow Frac Central
\input{../preamble/group_d_data_pick/pcky_var/CKY1000_mpoly.tex}
% Borrow Frac Northeast
\input{../preamble/group_d_data_pick/pcky_var/CKY15_msimu.tex}
% Informal Borrow Shares
\input{../preamble/group_d_data_pick/pcky_var/CKY15_msimu_ibbf_ne.tex}
\input{../preamble/group_d_data_pick/pcky_var/CKY15_msimu_ibbf_ce.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% CKY AMOUNTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% K Amount NE and CE and Y for CE both from SIMU 15
\input{../preamble/group_d_data_pick/pcky_var/CKY15_msimu_kne.tex}
\input{../preamble/group_d_data_pick/pcky_var/CKY15_msimu_kce.tex}
\input{../preamble/group_d_data_pick/pcky_var/CKY15_msimu_yce.tex}
\input{../preamble/group_d_data_pick/pcky_var/CKY1_msimu_cne_v2.tex}
% Y Amount NE from mpoly 100
% \input{../preamble/group_d_data_pick/pcky_var/CKY100_mpoly_yne.tex}
\input{../preamble/group_d_data_pick/pcky_var/CKY500_mpoly_yne_v2.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Probabilities Means
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Probabilities
\input{../preamble/group_d_data_pick/pcky_raw/PROB1.tex}
% replace prob confidence interval
\input{../preamble/group_d_data_pick/pcky_var/PROB500.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ACTUAL DATA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{../preamble/group_d_data_pick/pcky_data.tex}

\FITPRBKY{Model Fits}