### 2019-04-19 sections_combine/draft_main_blank.tex

\documentclass[12pt,english]{article}
\input{../fragments/preamble_main.tex}
\begin{document}

\title{Blank File}
\author{Fan Wang\thanks{Department of Economics, University of Houston (email: fwang26@uh.edu), \url{https://fanwangecon.github.io/research.html}}}
\maketitle

\input{../sections_blank/introconclude/abstract.tex}
\begin{abstract}
\singlespacing \abstractB
\end{abstract}
\vfill
\pagebreak{}

\section{Introduction}
\input{../sections_blank/introconclude/intro.tex}
\paragraph{Related Literature}
\input{../sections_blank/introconclude/literature.tex}

\section{Data}
\subsection{Data One}
\input{../sections_blank/data/data_one.tex}
\subsection{Data Two}
\input{../sections_blank/data/data_two.tex}
\subsection{Data Three}
\input{../sections_blank/data/data_two.tex}

\section{Model}
\subsection{Model One}
\input{../sections_blank/model/model_one.tex}
\subsection{Model Two}
\input{../sections_blank/model/model_two.tex}
\subsection{Model Three}
\input{../sections_blank/model/model_two.tex}

\section{Estimate}
\subsection{Estimation One}
\input{../sections_blank/esti/esti_one.tex}
\subsection{Estimation Two}
\input{../sections_blank/esti/esti_two.tex}
\subsection{Estimation Three}
\input{../sections_blank/esti/esti_two.tex}

\section{Conclusion}
\input{../sections_blank/introconclude/conclude.tex}

\pagebreak
\input{../fragments/cite/cite_end.tex}
\pagebreak

\appendix

\end{document}