### 2019-06-25 _other/pandoc/fansampleBACKDOCX.tex
pandoc converted to doc from updated tex

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Sandbox Testing Latex},
            pdfauthor={Fan Wang (University of Houston)},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{Sandbox Testing Latex}
\author{Fan Wang (University of Houston)}
\date{}

\begin{document}
\maketitle

\hypertarget{equation-test}{%
\section{Equation Test}\label{equation-test}}

Inline:
\(\sigma_{y} = \sqrt{\sum_{y}^{}P(Y = y) \cdot \left( y - \mu_{y} \right)^{2}}\)

\[\Omega^{\max} = \exp(z^{\max}) \cdot \left( \Lambda^{\max} \right)^{\alpha}\]

\[\sum_{i = 0}^{\infty}a_{i}x^{i}\]

\hypertarget{citation-testing}{%
\section{Citation Testing}\label{citation-testing}}

Becker and Tomes (1986) is a paper on Human Capital. Early childhood
health and economic circumstances could have lasting effects on adult
socioeconomic outcomes (Case, Fertig, and Paxson 2005).\footnote{"In
  this article, we develop a framework for analyzing the causal effects
  of interventions in the presence of latent factors that could affect
  outcomes, even in the absence of interventions" (Conti and Heckman
  2010).}

\hypertarget{url-test}{%
\section{URL Test}\label{url-test}}

\begin{itemize}
\item
  \href{http://fanwangecon.github.io}{Fan Wang Site}
\item
  \href{https://fanwangecon.github.io/CodeDynaAsset/}{Fan Wang Dynamic
  Asset Code Repository}
\end{itemize}

\hypertarget{pandoc}{%
\section{Pandoc}\label{pandoc}}

\hypertarget{install-and-convert-tex-to-word}{%
\subsection{Install and convert Tex to
Word}\label{install-and-convert-tex-to-word}}

For editing latex files in word to benefit from tracked changes.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  install pandoc: \href{https://pandoc.org/installing.html}{Pandoc Site
  Installation Link}
\item
  open up command line
\item
  \emph{cd into where tex file is}: cd
\item
  \emph{run this for tex to word}: pandoc -s fansample.tex -o
  fansample.docx
\end{enumerate}

\hypertarget{convert-from-tex-to-word-with-bibliography}{%
\subsection{Convert from Tex to Word with
bibliography}\label{convert-from-tex-to-word-with-bibliography}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \emph{run this for word to tex}: pandoc --bibliography=fansample.bib
  -o fansample.docx fansample.tex
\end{enumerate}

\hypertarget{convert-from-word-from-tex}{%
\subsection{Convert from Word from
Tex}\label{convert-from-word-from-tex}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \emph{run this for word to tex}: pandoc -s fansample.docx -o
  fansampleBACKDOCX.tex
\end{enumerate}

Becker, Gary S, and Nigel Tomes. 1986. ``Human Capital and the Rise and
Fall of Families.'' \emph{Journal of Labor Economics} 4 (3): S1--39.
\url{http://ideas.repec.org/a/ucp/jlabec/v4y1986i3ps1-39.html}.

Case, Anne, Angela Fertig, and Christina Paxson. 2005. ``The Lasting
Impact of Childhood Health and Circumstance.'' \emph{Journal of Health
Economics} 24 (2): 365--89.
\url{https://doi.org/10.1016/j.jhealeco.2004.09.008}.

Conti, Gabriella, and James J Heckman. 2010. ``Understanding the Early
Origins of the Education--Health Gradient.'' \emph{Perspectives on
Psychological Science} 5 (5): 585--605.
\url{https://doi.org/10.1177/1745691610383502}.

\end{document}