### 2019-09-12 _support/spacing/font.tex
spacing

\documentclass[12pt,english]{article}
\usepackage[utf8]{inputenc}

% url package
\usepackage{hyperref}

% Bibliography package
\usepackage[authordate, backend=biber]{biblatex-chicago}
\bibliography{fansample}

% Titling and Author
\title{Latex Font}
\author{Fan Wang (University of Houston)}
\date{\today}

\begin{document}

There are different font families in latex. \textbf{mathpazo} is a family type with nice math equations, but the texts and equations take up more space, potentially exceeding journals' submission limits.

\end{document}