### 2019-12-22 _other/tikz/tree/fs_tikz_gametree.tex

\documentclass{article}

\usepackage{blindtext}
\usepackage{multicol}
\usepackage{caption}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
\usepackage{geometry}
\geometry{
	a4paper,
	noheadfoot=true,
	left=1.0in,
	right=1.0in,
	top=1.0in,
	bottom=1.0in
}

% url package
\usepackage{hyperref}
\usepackage{subcaption}

% Titling and Author
\title{Latex Tikz Examples, Discrete and Continuous Strategy Trees}
\author{\href{https://fanwangecon.github.io/}{Fan Wang}\thanks{https://fanwangecon.github.io, repository: \href{https://fanwangecon.github.io/Tex4Econ/}{Tex4Econ}}}
\date{\today}

\begin{document}

\maketitle

% A.1 Node styles
\tikzset{
    % Two node styles for game trees: solid and hollow
    solid node/.style={circle,draw,inner sep=1.5,fill=black},
    hollow node/.style={circle,draw,inner sep=1.5,fill=white}
}

\begin{center}
\begin{tikzpicture}[scale=1.5,font=\footnotesize]

% A.2 Level Styles
    % Specify spacing for each level of the tree
    \tikzstyle{level 1}=[level distance=17mm,sibling distance=25mm]
    \tikzstyle{level 2}=[level distance=15mm,sibling distance=15mm]

% B Level 1 point
\node(0)[solid node,label=above:{$P1$}]{}

% C Level 2 Node (child nodes)
% C.1 First child node, left
child{
    % C.1.a Node line color
    [red]
    % C.1.b Node dot 
    node(1)[solid node, label=below:{red dot}]{}
    % C.1.c now go to edge
    edge from parent 
        % C.1.c Node along edge
        node[sloped, above, black, text width=3cm]{Minimum Choice showing up along this red line}
    }
% C.2 Middle Child node, do not show line, invisible
child{
    % Main node of level 2 middle node, same y-axis height as node left and right
    % Y-shift to move level 3 child node lower
    [purple] node(2)[solid node, xshift=30, yshift=-50, label=right:{hello}]{}
    % D.1 Left level 3 (from middle level 2)
    child{
        [black] 
        node[hollow node,label=below:{$(a,b)$}]{} 
        edge from parent 
            node[left]{$C$}
        }
    child{
        [black]
        node[hollow node,label=below:{$(c,d)$}]{} 
        edge from parent 
            node[right]{$D$}
        }
    edge from parent 
        %note that you need to adjust the yshift if you change the level distance
        node[left, black, xshift=-5, yshift=0, text width = 1mm]{$\alpha$\\$\beta$\\$\sigma$} 
    }
child{
    [black]
    node(3)[solid node, label=right:{hi there}, green]{}
    edge from parent 
        node[right, xshift=0, yshift=15, text width=2cm]{Some text for this edge}
    };
% information set
\draw[dashed, bend right] 
    (1) to (3);
\draw[dashed, bend left, line width=0.5mm, blue] 
    (0) to 
        node[right, text width=1.25cm, xshift=5, yshift=-10] 
            {\tiny{label right of blue dashed line}}
        (2);
\end{tikzpicture}
\end{center}

% \begin{istgame}[font=\scriptsize]
% \setistgrowdirection{east}
% \cntmdistance{20mm}{20mm}
% \cntmAistb{q_1=0}[at end,below]{q_1=1,000}[at end,above]
% \istrootcntmA(0){1}
%   \istbA[draw=none]
%   \endist
% \cntmAistb{q_2=0}[at end,below]{q_2=1,000}[at end,above]
% \istrootcntmA(1)(0-1)<[xshift=3pt]90>{2}
%   \istbA[draw=none]{}{\pi_1,\pi_2}
%   \endist
% \end{istgame}

\end{document}