### 2019-12-22 _other/pandoc/fansample.tex
Tex etc updates

\documentclass[12pt,english]{article}
\usepackage[utf8]{inputenc}

% url package
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,
    urlcolor=cyan,
}

% Bibliography package
\usepackage[authordate, backend=biber]{biblatex-chicago}
\bibliography{fansample}

% Titling and Author
\title{Latex Example, Pandoc Conversion}
\author{Fan Wang (University of Houston)}
\date{\today}

\begin{document}

\maketitle

\section{Equation Test}

Inline: $\sigma_y = \sqrt{ \sum_{y} P(Y=y) \cdot \left( y - \mu_y \right)^2}$

$$\Omega^{\max} = \exp(z^{\max}) \cdot \left(\Lambda^{\max}\right)^{\alpha}$$

\begin{equation}
\label{eq:1}
\sum_{i=0}^{\infty} a_i x^i
\end{equation}

\section{Citation Testing}

\textcite{becker_human_1986} is a paper on Human Capital. Early childhood health and economic circumstances could have lasting effects on adult socioeconomic outcomes \autocite{case_lasting_2005}.\footnote{"In this article, we develop a framework for analyzing the causal effects of interventions in the presence of latent factors that could affect outcomes, even in the absence of interventions" \autocite{conti_understanding_2010}.}

\section{URL Test}

\begin{itemize}
	\item \href{http://fanwangecon.github.io}{Fan Wang Site}
	\item \href{https://fanwangecon.github.io/CodeDynaAsset/}{Fan Wang Dynamic Asset Code Repository}
\end{itemize}

\pagebreak

\section{Pandoc}

\subsection{Install and convert Tex to Word}

For editing latex files in word to benefit from tracked changes.

\begin{enumerate}
	\item install pandoc: \href{https://pandoc.org/installing.html}{Pandoc Site Installation Link}
	\item open up command line
	\item \textit{cd into where tex file is}: cd \path{C:/Users/fan/Tex4Econ/_other/pandoc}
	\item \textit{run this for tex to word}: pandoc -s fansample.tex -o fansample.docx
\end{enumerate}

\subsection{Convert from Tex to Word with bibliography}

\begin{enumerate}
	\item \textit{run this for word to tex}: pandoc --bibliography=fansample.bib -o fansample.docx fansample.tex
\end{enumerate}

\subsection{Convert from Word from Tex}

\begin{enumerate}
	\item \textit{run this for word to tex}: pandoc -s fansample.docx -o fansampleBACKDOCX.tex
\end{enumerate}

\printbibliography

\end{document}