### 2019-12-22 nontex/install/linux/fn_ubuntu.Rmd
edit improve linux and windows installation files.

---
title: "Atom, Git, Github, etc Set-up for Linux"
output:
  pdf_document:
    df_print: paged
    number_sections: true
  html_notebook:
    number_sections: true
  word_document:
    number_sections: true
  html_document:
    number_sections: true
urlcolor: blue
always_allow_html: yes
---

Go back to [fan](http://fanwangecon.github.io/)'s [Miscellaneous Repository](http://fanwangecon.github.io/Tex4Econ/nontex/).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Atom Install

Can be directly installed from Ubuntu store.

```{bash}
# install atom
sudo apt-get install atom
```

- *personal access token*: d8a293d30ef620a8b27ab0ad6294cdd8942a0b6f
- *Gist id*: 754fa9ca57eb7ddc1ee89775f934b59f

# Git Install

1. install git
2. once rsa set up, sync repo

```{bash}
# install git
sudo apt-get install git
# generate repo directory and sync
mkdir ~/PyFan
cd ~/PyFan
git init
git config --global user.name "Fan Wang"
git config --global user.email wangfanbsg75@live.com
git remote add github git@github.com:FanWangEcon/PyFan.git
git pull github master
```

# Github and local repo

1. generate rsa
    - ssh-keygen -t rsa -C "wangfanbsg75@live.com", when prompted, do not enter "file in which save the key", when prompted for passphrase, enter "WHATEVERPASSWORDIS"
2. copy key
3. log on to github ssh section, generate new ssh rsa key

```{bash}
ssh-keygen -t rsa -C "wangfanbsg75@live.com"
cat ~/.ssh/id_rsa.pub
```

# Conda Install

conda install linux

1. wget to download from url to download folder
2. bash to install
3. follow instructions, type yes
4. source ~/.bashrc

```{bash}
# could be saved in current folder: pwd
# could be saved in download folder: ~/Downloads
wget https://repo.anaconda.com/archive/Anaconda3-2019.10-Linux-x86_64.sh
# install file
bash "Anaconda3-2019.10-Linux-x86_64.sh"
# refresh
source ~/.bashrc
# show all installed packages under current envir
conda list
# see where key files are installed
which python
which conda
# use conda-forge as main channel and update all
conda config --add channels conda-forge
conda update --all
```

## Install PyCharm

PyCharm can be installed from Ubuntu apps.

```{bash}
<!-- for Debian -->
sudo tar xfz pycharm-*.tar.gz -C /opt/
cd /opt/pycharm-*/bin
./pycharm.sh
```

# R install

R could be installed first as below. Or follow the [instructions on this page](https://fanwangecon.github.io/Tex4Econ/nontex/install/windows/fn_installations.html) to install from conda an environment for R, with associated R-studio.

```{bash}
apt-get update
apt-get install r-base r-base-dev
```