### 2019-12-29 nontex/install/linux/fn_vim.Rmd
start vim set-up and knowledge file

---
title: "Example and Tutorial for Vim Editor (Linux)"
output:
  pdf_document:
    df_print: paged
    number_sections: true
  html_notebook:
    number_sections: true
  word_document:
    number_sections: true
  html_document:
    number_sections: true
urlcolor: blue
always_allow_html: yes
---

Go back to [fan](http://fanwangecon.github.io/)'s [Tex4Econ and Miscellaneous Repository](http://fanwangecon.github.io/Tex4Econ/).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Vim Installations

Vim ships with Debian.

Additional packages to install:

```{bash}
vim
```

Also install neo-vim

```{bash}
sudo apt-get install neovim
```

## Plug-ins

Follow the [instructions here](https://github.com/junegunn/vim-plug/wiki/tutorial) for installations. Two steps:

1. generate ~/.vim and ~/.vimrc folders
2. curl plug.vim to ~/.vim/autoload
3. update ~/.vimrc file with ininstructions regarding which files to install.
4. open vim, type: *:PlugInstall* or *:PlugUpdate*, should start seeing updating packages.
5. type: *:scriptnames* to see loaded plugins
6. just comment out or delete plug lines for packages no longer needed.
7. how to install a particular package? See [vimawesome](https://vimawesome.com/plugin/onedark-vim)

```{bash}
# curl plug.vim
curl -fLo ~/.vim/autoload/plug.vim --create-dirs \
    https://raw.githubusercontent.com/junegunn/vim-plug/master/plug.vim

# create vimrc file
cd $HOME
vim .vimrc

# paste into file the text below.
```

Example .vimrc file with only one package one dark atom visual for vim:

```{bash}
if empty(glob('~/.vim/autoload/plug.vim'))
        silent !curl -fLo ~/.vim/autoload/plug.vim --create-dirs
                \ https://raw.githubusercontent.com/junegunn/vim-plug/master/plug.vim
        autocmd VimEnter * PlugInstall --sync | source $MYVIMRC
endif

" Plugins will be downloaded under the specified directory.
call plug#begin('~/.vim/plugged')

" Declare the list of plugins.
Plug 'joshdick/onedark.vim'

" List ends here. Plugins become visible to Vim after this call.
call plug#end()

"""""""""""""""""""""""""""""""""""""""""""""""""
" Settings
"""""""""""""""""""""""""""""""""""""""""""""""""
syntax on
colorscheme onedark
```

# Vim Basics

```{bash}
# to open vim
vim -v
# to exit vim
:q!
# open up file from google drive
vim "/mnt/chromeos/GoogleDrive/MyDrive/Reading Papers/Reading/Referee/pa_2019_oct_journal/jor_12345_comments.Rmd"
# start editing, type i to enter insert mode
i
# when editing is done, to exit and save, write and quit
# esc first
:wq
```

# Vim Cursor Movements

Move cursor to the beginning of a line? This should not be done in edit mode, esc, use movement commands, then back to insert mode. There are six most basic movements, *wbjk()0$*:

```{bash}
# forward one word
w
# back one word to begin
b
# back to word to end
e

# forward one line, achieved by also pressing downward arrow
j
# backward one line, achieved by pressing upward arrow
k
# left arrow
h
# right arrow
l

# Jump forward one Sentence
)
# Jump backward one Sentence
(

# Begin of line
0
# End of Line
$

# Begin of File
gg
# End of File
G
```

In combination, the basic commands can be used in combination

```{bash}
# move to the beginning of the next line
j0
# move to the end of the next line
j$
```

# Vim Editing

Delete a line.

```{bash}
# select text with cursor, then d for delete
d
```

# Very Convinent Functions

Some functionalities take much more efforts to complete in many other editors, but are very easy in vim:

```{bash}
# combine lines: bunch of lines, with carriage break, one block of text, reflow/join
# 1. select block
Vjjjj
# 2. reflow/join
J
```

## Combo Operations

1. Delete:
    - delete current word and go into insert mode: *ciw*
    - Delete current word: *dw*
    - Delete from cursor until end of sentence: *dis*
    - Delete from cursor to end of line: *d$* or *D*
    - Delete until end of paragraph: *dip*

# Vim Errors

Vim seems frozen (hard for vim to freeze, it is not really frozen). This is likely due to pressing *ctrl + s*, then need to press *ctrl + q*.

vim error: "E297: Write error in swap file"

```
# Trying to save
:w
# E297 can not save, override changes elswhere, and force save what is on screen.
:e!
```

# Vim Set-Up

Use [one-dark atom theme](https://github.com/joshdick/onedark.vim), otherwise color hard to read.

```{bash}

```