### 2020-01-03 nontex/install/windows/fn_installations.tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={New Computer Data-Science Set-up: Python (anaconda, pycharm), R (r-studio) and other Programs Installations},
            colorlinks=true,
            linkcolor=Maroon,
            citecolor=Blue,
            urlcolor=blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{New Computer Data-Science Set-up: Python (anaconda, pycharm), R
(r-studio) and other Programs Installations}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\begin{document}
\maketitle

Go back to \href{http://fanwangecon.github.io/}{fan}'s
\href{http://fanwangecon.github.io/Tex4Econ/}{Tex4Econ and Miscellaneous
Repository}.

\hypertarget{objective}{%
\section{Objective}\label{objective}}

For the following types of programming tasks, install relevant programs,
that are cross platform compatible. Test at the end that example
programs work.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Python

  \begin{itemize}
  \tightlist
  \item
    Anaconda Dist
  \item
    Pycharm
  \item
    atom hydrogen
  \end{itemize}
\item
  R

  \begin{itemize}
  \tightlist
  \item
    R
  \item
    Rstudio
  \item
    atom hydrogen
  \end{itemize}
\item
  Matlab

  \begin{itemize}
  \tightlist
  \item
    Matlab
  \end{itemize}
\item
  Latex and PDF

  \begin{itemize}
  \tightlist
  \item
    atom latex
  \item
    texstudio
  \item
    texlive
  \item
    okular pdf
  \end{itemize}
\end{enumerate}

Test at the end:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Do sample \emph{py} files work?

  \begin{itemize}
  \tightlist
  \item
    work from command line?
  \item
    work from atom (hydrogen)
  \end{itemize}
\item
  Do sample \emph{rmd} files work?

  \begin{itemize}
  \tightlist
  \item
    work from r-studio
  \item
    work from atom (hydrogen), r file
  \end{itemize}
\item
  Do tex templates compile?

  \begin{itemize}
  \tightlist
  \item
    tikz files
  \item
    template files with bib
  \end{itemize}
\end{enumerate}

\emph{note}: Github people seem to have slowed down updating Atom since
microsoft acquisition

\hypertarget{conda}{%
\section{Conda}\label{conda}}

Use conda across platforms, so that locally on windows and ubuntu and
remotely on aws, can have the same software setup environment.

Search for Anaconda Prompt, right click, choose run as administrator.

Check software versions.

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{conda}\NormalTok{ list anaconda}
\ExtensionTok{python}\NormalTok{ -V}
\end{Highlighting}
\end{Shaded}

\hypertarget{first-time-conda-install}{%
\subsection{First Time Conda Install}\label{first-time-conda-install}}

Download and install for all users. Afterwards, Anaconda does not
automatically get added to Windows Path. Need to use Anaconda Prompt to
access programs. To access Anaconda packages from windows prompt, from
git bash, from R, etc, need to Add Anaconda to Windows Path

To install conda in Debian, follow
\href{https://fanwangecon.github.io/Tex4Econ/nontex/install/linux/fn_ubuntu.html}{these
instructions}.

See Anaconda Installation Directory, in anaconda prompt:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# To remove conda Fully}
\FunctionTok{rm}\NormalTok{ -rf ~/anaconda3}

\ExtensionTok{where}\NormalTok{ anaconda}
\CommentTok{# C:/ProgramData/Anaconda3/Scripts/anaconda.exe}
\ExtensionTok{where}\NormalTok{ python}
\CommentTok{# C:/ProgramData/Anaconda3/python.exe}
\ExtensionTok{where}\NormalTok{ jupyter}
\CommentTok{# C:/ProgramData/Anaconda3/Scripts/jupyter.exe}
\ExtensionTok{where}\NormalTok{ jupyter-kernelspec}
\CommentTok{# if R installed already}
\ExtensionTok{where}\NormalTok{ r}

\CommentTok{###############################}
\CommentTok{# Add these to Windows PATH:}
\CommentTok{###############################}
\CommentTok{# C:\textbackslash{}ProgramData\textbackslash{}Anaconda3\textbackslash{}Scripts}
\CommentTok{# C:\textbackslash{}ProgramData\textbackslash{}Anaconda3}
\end{Highlighting}
\end{Shaded}

To Add Anaconda to Path, In Windows 1. search for: Environment Variables
2. Edit Environment Variables 3. Add new to Path (lower half): -
C:/ProgramData/Anaconda3/Scripts/ - C:/ProgramData/Anaconda3/ 4. Now
open up regular windows command Prompt, Type in: - conda --version -
also Close and Open up Git Bash: conda --version

\hypertarget{conda-update}{%
\subsection{Conda Update}\label{conda-update}}

Open up Anaconda Navigator, it will update navigator automatically. If
there are errors, might have to clean first.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# if there are bugs}
\CommentTok{# conda clean --packages}
\CommentTok{# use conda-forge as main channel, more updated packages}
\ExtensionTok{conda}\NormalTok{ config --get channels}
\ExtensionTok{conda}\NormalTok{ config --add channels conda-forge}
\ExtensionTok{conda}\NormalTok{ config --get channels}
\CommentTok{# remove conda-forge from channels (do so for main env install)}
\ExtensionTok{conda}\NormalTok{ config --remove channels conda-forge}
\CommentTok{# normal update}
\ExtensionTok{conda}\NormalTok{ update --all}

\CommentTok{# install additional packages}
\ExtensionTok{conda}\NormalTok{ install -y statsmodels datashape seaborn}
\ExtensionTok{conda}\NormalTok{ install -c conda-forge -y interpolation awscli}
\ExtensionTok{conda}\NormalTok{ install -c anaconda -y boto3}
\end{Highlighting}
\end{Shaded}

\hypertarget{conda-test-python}{%
\subsection{Conda Test Python}\label{conda-test-python}}

Fist test python under command line:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# windows}
\ExtensionTok{python} \StringTok{"C:/Users/fan/PyFan/ProjectSupport/Testing/Numpy/Functions.py"}
\ExtensionTok{python} \StringTok{"C:/Users/fan/PyFan/ProjectSupport/graph/subplot.py"}
\CommentTok{# linux}
\ExtensionTok{python}\NormalTok{ ~/PyFan/ProjectSupport/Testing/Numpy/Functions.py}
\ExtensionTok{python}\NormalTok{ ~/PyFan/ProjectSupport/graph/subplot.py}
\end{Highlighting}
\end{Shaded}

\hypertarget{jupyter-and-atom-hydrogen}{%
\subsection{Jupyter and Atom Hydrogen}\label{jupyter-and-atom-hydrogen}}

For both Jupyter as well as Atom, always open from command prompt, open
from \emph{Anaconda Prompt} with admin rights.

First, open up \emph{Jupyter Notebook}, does the python kernel work? If
does not, uninstall and re-install Anaconda.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# start jupyter lab, from prompt}
\ExtensionTok{jupyter}\NormalTok{ lab}
\end{Highlighting}
\end{Shaded}

Now test python within atom:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  install hydrogen in atom. Follow
  \href{http://nteract.io/atom}{instructions}.

  \begin{itemize}
  \tightlist
  \item
    terminal in windows or linux after installation: \emph{apm install
    hydrogen}
  \item
    apm is the atom package manager.
  \end{itemize}
\item
  Pull/Sync with Git and open mulitple projects in atom, follow
  \href{https://fanwangecon.github.io/Tex4Econ/nontex/install/git/fs_git_setup.html}{guide
  here}.
\item
  go to the programs above and try to run.
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Once atom is installed, can use apm to install packages}
\FunctionTok{apm}\NormalTok{ install hydrogen}
\FunctionTok{apm}\NormalTok{ install project-manager}
\FunctionTok{apm}\NormalTok{ install sublime-style-column-selection}
\FunctionTok{apm}\NormalTok{ install minimap}

\CommentTok{# vim distraction free editing}
\FunctionTok{apm}\NormalTok{ install vim-mode-plus}
\FunctionTok{apm}\NormalTok{ install zen}

\CommentTok{# Inside Conda open up atom}
\ExtensionTok{atom}

\CommentTok{# open multiple projects from atom at the same time, after repos synced}
\ExtensionTok{atom}\NormalTok{ ~/fanwangecon.github.io ~/Pyfan ~/Teaching ~/Tex4Econ ~/R4Econ ~/M4Econ ~/Py4Econ}
\CommentTok{# open up any py from from within, try 1+1, does it work?}
\CommentTok{# also use script, try ctrl + shift + b, run whole file (not hydrogen)}
\end{Highlighting}
\end{Shaded}

\hypertarget{r-installation}{%
\section{R Installation}\label{r-installation}}

Try to install R from within Conda. The procedure here is the same for
windows as well as linux machines.

\hypertarget{fully-uninstall-r}{%
\subsection{Fully Uninstall r}\label{fully-uninstall-r}}

Open up RStudio, and see where paths are. also open up R from command
prompt, see where paths are.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{.libPath}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

Open up install and uninstall programs, find R and Rstudio, and
uninstall them. Then delete all files found path paths shown by
\emph{.libPath()}, looking at these: -
\emph{C:/Users/fan/Documents/R/win-library/3.6} - \emph{C:/Program
Files/R/R-3.6.1/library}

For Linux:

\begin{itemize}
\tightlist
\item
  if used apt-get to install, R installed under main user drive,
  deactivate conda, and check which conda
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Exit Conda}
\ExtensionTok{conda}\NormalTok{ deactivate}
\CommentTok{# where is R installed outside of Conda}
\FunctionTok{which}\NormalTok{ R}
\CommentTok{# /usr/bin/R}
\CommentTok{# To remove all}
\FunctionTok{sudo}\NormalTok{ apt-get remove r-base}
\FunctionTok{sudo}\NormalTok{ apt-get remove r-base-core}

\CommentTok{# Inside Conda base}
\ExtensionTok{conda}\NormalTok{ activate}
\CommentTok{# Conda r_env}
\ExtensionTok{conda}\NormalTok{ activate r_env}
\CommentTok{# Where is it installed?}
\FunctionTok{which}\NormalTok{ R}
\CommentTok{# /home/wangfanbsg75/anaconda3/bin/R}
\ExtensionTok{conda}\NormalTok{ uninstall r-base}
\end{Highlighting}
\end{Shaded}

\hypertarget{install-r-outside-of-conda}{%
\subsection{Install R (outside of
Conda)}\label{install-r-outside-of-conda}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \href{https://cloud.r-project.org/}{download R}

  \begin{itemize}
  \tightlist
  \item
    for debian:
    \href{https://cran.r-project.org/bin/linux/debian/}{Johannes Ranke}
  \end{itemize}
\item
  \href{https://rstudio.com/products/rstudio/download/}{download
  R-studio}
\item
  Open R-studio and auto-detect R
\item
  Install packages
\end{enumerate}

note that installation will not work if the machine's version of R is
not up to date, this could be the case with default R versions for some
linux installations. For Debian for example, need to update the
\emph{source.list} to include sources that have more recent versions of
R.

Additionally, need to install \emph{libcurl4-openssl-dev} and
\emph{libssl-dev} from command prompt first inside Debian.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# If inside Debian, due to some tidymodel install issues, do this`}
\ExtensionTok{conda}\NormalTok{ install -c r r}
\ExtensionTok{conda}\NormalTok{ install -c r r-essentials}
\ExtensionTok{conda}\NormalTok{ install -c r r-igraph}
\ExtensionTok{conda}\NormalTok{ install -c conda-forge igraph}

\CommentTok{# Go to get latesdebian latest r sources.list}
\FunctionTok{cat}\NormalTok{ /etc/apt/sources.list}
\CommentTok{# Install this First (should already be installed)}
\FunctionTok{sudo}\NormalTok{ apt install dirmngr}

\CommentTok{# Debian R is maintained by Johannes Ranke, copied from https://cran.r-project.org/bin/linux/debian/:}
\ExtensionTok{apt-key}\NormalTok{ adv --keyserver keys.gnupg.net --recv-key }\StringTok{'E19F5F87128899B192B1A2C2AD5F960A256A04AF'}
\CommentTok{# Add to source.list, for debian stretch (9)}
\CommentTok{# sudo su added for security issue as super-user}
\FunctionTok{sudo}\NormalTok{ su -c }\StringTok{"sudo echo 'deb http://cloud.r-project.org/bin/linux/debian stretch-cran35/' >> /etc/apt/sources.list"}
\CommentTok{# if added wrong lines, delete 3rd line}
\FunctionTok{sudo}\NormalTok{ sed }\StringTok{'3d'}\NormalTok{ /etc/apt/sources.list}

\CommentTok{# Update and Install R, should say updated from cloud.r}
\FunctionTok{sudo}\NormalTok{ apt-get update}
\FunctionTok{sudo}\NormalTok{ apt-get install r-base r-base-dev}

\CommentTok{# Also install these, otherwise r-packages do not install}
\CommentTok{# libxml2 seems need for tidymodels}
\FunctionTok{sudo}\NormalTok{ apt-get install libcurl4-openssl-dev}
\FunctionTok{sudo}\NormalTok{ apt-get install libssl-dev}
\FunctionTok{sudo}\NormalTok{ apt-get install libxml2-dev}
\end{Highlighting}
\end{Shaded}

Which R?

\begin{itemize}
\tightlist
\item
  \emph{export RSTUDIO\_WHICH\_R=/usr/bin/R}
\item
  \emph{export RSTUDIO\_WHICH\_R=/home/wangfanbsg75/anaconda3/bin/R}
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Can enter R from Command Prompt Conda ENV, much faster than from r-studio}
\CommentTok{# Install R-tools, RTOOLS is for WINDOWS}
\KeywordTok{install.packages}\NormalTok{(}\StringTok{"installr"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"installr"}\NormalTok{)}
\KeywordTok{install.Rtools}\NormalTok{()}
\CommentTok{# c:/Rtools/bin;}
\CommentTok{# c:/Rtools/mingw_32/bin;}
\CommentTok{# c:/Rtools/mingw_64/bin;}

\CommentTok{# main packagevps}
\CommentTok{# tidymodel installation on linux difficult}
\KeywordTok{install.packages}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"tidyverse"}\NormalTok{, }\StringTok{"tidymodels"}\NormalTok{, }\StringTok{"tidyr"}\NormalTok{))}
\CommentTok{# development packages}
\KeywordTok{install.packages}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"devtools"}\NormalTok{, }\StringTok{"IRkernel"}\NormalTok{, }\StringTok{"pkgdown"}\NormalTok{, }\StringTok{"roxygen2"}\NormalTok{, }\StringTok{"kableExtra"}\NormalTok{))}
\CommentTok{# other packages}
\KeywordTok{install.packages}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"AER"}\NormalTok{, }\StringTok{"minpack.lm"}\NormalTok{, }\StringTok{"knitr"}\NormalTok{, }\StringTok{"matlab"}\NormalTok{))}

\CommentTok{# Install my package, need to prepare package to work for ubuntu}
\CommentTok{# if install does not work, load locally}
\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"fanwangecon/R4Econ"}\NormalTok{)}

\CommentTok{# Kernel}
\KeywordTok{install.packages}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{'repr'}\NormalTok{, }\StringTok{'IRdisplay'}\NormalTok{, }\StringTok{'evaluate'}\NormalTok{, }\StringTok{'crayon'}\NormalTok{, }\StringTok{'pbdZMQ'}\NormalTok{, }\StringTok{'devtools'}\NormalTok{, }\StringTok{'uuid'}\NormalTok{, }\StringTok{'digest'}\NormalTok{))}
\NormalTok{IRkernel}\OperatorTok{::}\KeywordTok{installspec}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\hypertarget{set-up-r-environment-in-conda}{%
\subsection{Set up R-Environment in
Conda}\label{set-up-r-environment-in-conda}}

Conda environments are language agnostic. pip is a package manager for
Python. venv is an environment manager for Python. conda is both a
package and environment manager and is language agnostic. Inside
Anaconda Prompt:

Key packages are generally available in conda's default channel
(official distribution) and also the more frequently updated conda-forge
channel. Prioritize conda-forge to get the latest packages. after adding
\emph{conda-forge} to channels, that will be prioritized, so when
creating a new environment, conda will install first from conda-forge if
package exists there. Try the \emph{r\_env} generation line with and
without first adding \emph{conda-forge} to channel, and see that the
packages installed will have different versions, reflecting versions in
the default channel and in the \emph{conda-forge} channels.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# outside of conda, install some depencies}
\FunctionTok{sudo}\NormalTok{ apt-get install libcurl4-openssl-dev}
\FunctionTok{sudo}\NormalTok{ apt-get install libssl-dev}
\FunctionTok{sudo}\NormalTok{ apt-get install libxml2-dev}

\CommentTok{# install key r packages within a r-environment}
\ExtensionTok{conda}\NormalTok{ env remove -n r_env}
\CommentTok{# start empty environment}
\ExtensionTok{conda}\NormalTok{ create -n r_env}
\CommentTok{# activate env and check, nothing yet}
\ExtensionTok{conda}\NormalTok{ activate r_env}
\ExtensionTok{conda}\NormalTok{ list}

\CommentTok{# check channel and add env specific channel}
\ExtensionTok{conda}\NormalTok{ config --get channels}
\CommentTok{# to get latest r, use conda-forge, which has more recent Rs}
\CommentTok{# conda config --env --add channels conda-forge}
\CommentTok{# channels for base}
\ExtensionTok{conda}\NormalTok{ config --get channels}
\CommentTok{# channels for cur env}
\ExtensionTok{conda}\NormalTok{ config --get channels --env}

\CommentTok{# install files}
\ExtensionTok{conda}\NormalTok{ install -n r_env -c r r r-essentials r-tidyverse r-tidymodels -y}

\CommentTok{# trouble with igraph installation, install here from conda}
\CommentTok{# conda install -n r_env r-essentials r-base r-tidyr r-tidyverse r-devtools r-irkernel r-pkgdown r-roxygen2}
\ExtensionTok{conda}\NormalTok{ install -n r_env -c r r-igraph}
\ExtensionTok{conda}\NormalTok{ install -n r_env -c conda-forge igraph}
\ExtensionTok{conda}\NormalTok{ install -n r_env -c r r-rstantools r-rstan r-rstanarm -y}

\CommentTok{# see all installed environments}
\ExtensionTok{conda}\NormalTok{ env list}
\CommentTok{# activate an environment to use it, if in base, r does not exist, isolated in r_env}
\ExtensionTok{conda}\NormalTok{ activate r_env}
\CommentTok{# see packages in environment}
\ExtensionTok{conda}\NormalTok{ list}
\CommentTok{# to quit}
\ExtensionTok{conda}\NormalTok{ deactivate}
\end{Highlighting}
\end{Shaded}

\begin{itemize}
\tightlist
\item
  \textbf{Location}: Installed environments are shown appear in the
  \emph{envs} folder of \emph{C:/ProgramData/Anaconda3/envs/}. Can be
  easily deleted without disrupting the main base installation for
  conda. This is very important. During installation, could easily run
  into problems, and need to re-install. Much better to only re-install
  a folder. Note that when R is installed in \emph{envs}, it will not
  show up under add or remove programs for uninstallation from there,
  has to be uninstalled, deleted directly here.
\item
  \textbf{Use only in Env}: Note that if we installed r inside r\_env,
  if we type \emph{r} under the base environment, we can not enter r. We
  can only enter r within r\_env. With PATH properly set-up, this
  happens under Conda Prompt, Windows Prompt, etc.
\item
  \textbf{Multiple Pythons}: With the above installation for R, one
  benefit is that if R requires a different version of Python than what
  the base environment uses, the r\_env could have a different python
  version. So type \emph{python} in the base environment as well as
  inside the \emph{r\_env}. This also means potentially there could be
  duplicated installations I think. Look at the python versions under
  \emph{conda list} in the base and in the r-environment.
\item
  \textbf{File Size}: Note that this creates a large installation
  folder, \emph{C:/ProgramData/Anaconda3/envs/r\_env/}, without
  additional packages, is 1.3 GB in size
\end{itemize}

\hypertarget{install-additional-packages-inside-the-r-environment}{%
\subsection{Install Additional Packages inside the
R-environment}\label{install-additional-packages-inside-the-r-environment}}

These are packages that I use, these should be installed inside the
environment. Or they could be installed later from inside r-studio.
Installing inside r-studio is a lot better.

\emph{tidymodel} has installation issues sometimes.

In \emph{r\_env}, type in \emph{R}, and then install from inside
terminal's R:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# enter into env}
\ExtensionTok{conda}\NormalTok{ activate r_env}
\CommentTok{# enter R or type in rstudio}
\CommentTok{# sicne R is only install inside r_env, has to enter rstudio from r_env}
\ExtensionTok{R}
\ExtensionTok{rstudio}

\CommentTok{# main packagevps}
\CommentTok{# tidymodel installation on linux difficult}
\ExtensionTok{install.packages}\NormalTok{(c(}\StringTok{"rstantools"}\NormalTok{, }\StringTok{"rstan"}\NormalTok{, }\StringTok{"rstanarm"}\NormalTok{))}
\ExtensionTok{install.packages}\NormalTok{(c(}\StringTok{"tidyr"}\NormalTok{, }\StringTok{"tidyverse"}\NormalTok{, }\StringTok{"tidymodels"}\NormalTok{))}

\CommentTok{# tidymodels}
\CommentTok{# development packages}
\ExtensionTok{install.packages}\NormalTok{(c(}\StringTok{"devtools"}\NormalTok{, }\StringTok{"IRkernel"}\NormalTok{, }\StringTok{"pkgdown"}\NormalTok{, }\StringTok{"roxygen2"}\NormalTok{, }\StringTok{"kableExtra"}\NormalTok{))}
\CommentTok{# other packages}
\ExtensionTok{install.packages}\NormalTok{(c(}\StringTok{"AER"}\NormalTok{, }\StringTok{"minpack.lm"}\NormalTok{, }\StringTok{"knitr"}\NormalTok{, }\StringTok{"matlab"}\NormalTok{))}

\CommentTok{# Install my package, need to prepare package to work for ubuntu}
\CommentTok{# if install does not work, load locally}
\ExtensionTok{devtools}\NormalTok{::install_github(}\StringTok{"fanwangecon/R4Econ"}\NormalTok{)}

\CommentTok{# Kernel}
\ExtensionTok{install.packages}\NormalTok{(c(}\StringTok{'repr'}\NormalTok{, }\StringTok{'IRdisplay'}\NormalTok{, }\StringTok{'evaluate'}\NormalTok{, }\StringTok{'crayon'}\NormalTok{, }\StringTok{'pbdZMQ'}\NormalTok{, }\StringTok{'devtools'}\NormalTok{, }\StringTok{'uuid'}\NormalTok{, }\StringTok{'digest'}\NormalTok{))}
\FunctionTok{IRkernel::installspec()}
\end{Highlighting}
\end{Shaded}

Install additional packages from inside R. The devtools install will
install packages inside \emph{envs/r\_env/Lib/R}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{#######################################}
\CommentTok{### Own Package}
\CommentTok{#######################################}
\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"fanwangecon/R4Econ"}\NormalTok{)}
\CommentTok{# install_dev("cli")}
\end{Highlighting}
\end{Shaded}

\hypertarget{r-studio-set-up-download-from-rstudio}{%
\subsection{R-Studio Set-up Download from
Rstudio}\label{r-studio-set-up-download-from-rstudio}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Download R-studio, and install, as normal (not from conda, directly
  from rstudio website latest windows version)
\item
  Open up anaconda prompt, enter \emph{r\_env} environment:
  \emph{activate r\_env}
\item
  cd into rstudio exe file folder: cd ``C:/Program Files/RStudio/bin''
\item
  start r-studio from inside \emph{r\_env}: rstudio.exe
\item
  inside r-studio, check: \emph{.libPaths()}

  \begin{itemize}
  \tightlist
  \item
    ``C:/ProgramData/Anaconda3/envs/r\_env/Lib/R/library''
  \end{itemize}
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{activate}\NormalTok{ r_env}
\BuiltInTok{cd} \StringTok{"C:/Program Files/RStudio/bin"}
\ExtensionTok{rstudio.exe}

\BuiltInTok{cd} \StringTok{"C:/ProgramData/Anaconda3/envs/r_env/Lib/R/library"}
\StringTok{"C:/Program Files/RStudio/bin/rstudio.exe"}
\end{Highlighting}
\end{Shaded}

For linux, note that rstudio offers Ubuntu as well as Debian versions
that can be installed with wget:

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{<}\NormalTok{!}\ExtensionTok{--}\NormalTok{ for ubuntu --}\OperatorTok{>}
\FunctionTok{wget} \StringTok{"https://download1.rstudio.org/desktop/bionic/amd64/rstudio-1.2.5033-amd64.deb"}
\OperatorTok{<}\NormalTok{!}\ExtensionTok{--}\NormalTok{ for debian --}\OperatorTok{>}
\FunctionTok{sudo}\NormalTok{ apt install ./rstudio-1.2.5033-amd64.deb}
\FunctionTok{wget} \StringTok{"https://download1.rstudio.org/desktop/debian9/x86_64/rstudio-1.2.5033-amd64.deb"}
\FunctionTok{sudo}\NormalTok{ apt install ./rstudio-1.2.5033-amd64.deb}
\ExtensionTok{conda}\NormalTok{ install -c r rstudio}
\BuiltInTok{cd} \StringTok{"C:/Program Files/RStudio/bin"}
\ExtensionTok{rstudio.exe}

\BuiltInTok{cd} \StringTok{"C:/ProgramData/Anaconda3/envs/r_env/Lib/R/library"}
\StringTok{"C:/Program Files/RStudio/bin/rstudio.exe"}
\end{Highlighting}
\end{Shaded}

\hypertarget{install-packages-from-r-studio}{%
\subsection{Install Packages from
R-Studio}\label{install-packages-from-r-studio}}

Install addition files inside Rstudio. Easier to debug potentially.
Install rtools as described
\href{https://thecoatlessprofessor.com/programming/cpp/installing-rtools-for-compiled-code-via-rcpp/}{here}:
\href{https://cran.r-project.org/bin/windows/Rtools/}{rtools download}.
Alternatively:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{install.packages}\NormalTok{(}\StringTok{"installr"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"installr"}\NormalTok{)}
\CommentTok{# Note that this will generate several pop-up windows}
\KeywordTok{install.Rtools}\NormalTok{()}
\CommentTok{# choose to only install 64 bit toolchain}
\CommentTok{# this installs g++ which is needed for rstanarm}
\CommentTok{# Add these below to windows path}
\CommentTok{# c:/Rtools/bin;}
\CommentTok{# c:/Rtools/mingw_64/bin;}

\ControlFlowTok{if}\NormalTok{ (}\OperatorTok{!}\KeywordTok{require}\NormalTok{(devtools)) \{}
  \KeywordTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{)}
  \KeywordTok{library}\NormalTok{(devtools)}
\NormalTok{\}}
\KeywordTok{library}\NormalTok{(devtools)}
\KeywordTok{install_github}\NormalTok{(}\StringTok{"stan-dev/rstanarm"}\NormalTok{, }\DataTypeTok{args =} \StringTok{"--preclean"}\NormalTok{)}

\KeywordTok{setwd}\NormalTok{(}\StringTok{'C:/ProgramData/Anaconda3/envs/r_env/Lib/R/library'}\NormalTok{)}

\KeywordTok{install.packages}\NormalTok{(}\StringTok{"stan-dev/rstanarm"}\NormalTok{, }\DataTypeTok{dependencies=}\OtherTok{TRUE}\NormalTok{, }\DataTypeTok{INSTALL_opts =} \KeywordTok{c}\NormalTok{(}\StringTok{'--no-lock'}\NormalTok{))}

\CommentTok{# other packages}
\KeywordTok{install.packages}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"tidymodels"}\NormalTok{, }\StringTok{"AER"}\NormalTok{, }\StringTok{"minpack.lm"}\NormalTok{, }\StringTok{"knitr"}\NormalTok{, }\StringTok{"kableExtra"}\NormalTok{, }\StringTok{"matlab"}\NormalTok{))}

\CommentTok{# for r-kernel to work in Atom}
\KeywordTok{install.packages}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{'repr'}\NormalTok{, }\StringTok{'IRdisplay'}\NormalTok{, }\StringTok{'evaluate'}\NormalTok{, }\StringTok{'crayon'}\NormalTok{, }\StringTok{'pbdZMQ'}\NormalTok{, }\StringTok{'devtools'}\NormalTok{, }\StringTok{'uuid'}\NormalTok{, }\StringTok{'digest'}\NormalTok{))}
\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{'IRkernel/IRkernel'}\NormalTok{)}
\NormalTok{IRkernel}\OperatorTok{::}\KeywordTok{installspec}\NormalTok{()}

\CommentTok{# Install my package}
\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"fanwangecon/R4Econ"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{test-files}{%
\subsubsection{Test Files}\label{test-files}}

Test the following file to see if we can execute a R file. Do it inside
\emph{r\_env} and inside a \emph{r} session.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# A simple file with summary statistics using tidyverse}
\KeywordTok{source}\NormalTok{(}\StringTok{'C:/Users/fan/R4Econ/summarize/dist/fst_hist_onevar.R'}\NormalTok{)}
\CommentTok{# Another simple file with summary statistics using tidyverse}
\KeywordTok{source}\NormalTok{(}\StringTok{'C:/Users/fan/R4Econ/support/tibble/fs_tib_basics.R'}\NormalTok{)}
\CommentTok{# A file involving estimation}
\KeywordTok{source}\NormalTok{(}\StringTok{'C:/Users/fan/R4Econ/optimization/cesloglin/fst_ces_plan_linlog.R'}\NormalTok{)}

\CommentTok{# C:/Users/fan/R4Econ/summarize/dist/fst_hist_onevar.Rmd}
\CommentTok{# C:/Users/fan/R4Econ/support/tibble/fs_tib_basics.Rmd}
\CommentTok{# C:/Users/fan/R4Econ/optimization/cesloglin/fst_ces_plan_linlog.Rmd}
\end{Highlighting}
\end{Shaded}

\hypertarget{install-latex}{%
\section{Install Latex}\label{install-latex}}

Use a combination of overleaf and local compile. The goal is to have the
same installation set-up for windows as well as linux. And hopefully,
locally compilable file also can be compiled on overleaf (only use
packages also available on overleaf).

There are several pieces of things to install:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  latex compiler: pdflatex, etc
\item
  tex distributions: MikTex, texlive

  \begin{itemize}
  \tightlist
  \item
    these folders, once various packages are installed, could be very
    large, many GB
  \end{itemize}
\item
  latex packages from distributions
\item
  editor (gui): Atom, Sublime text, texStudio, etc
\end{enumerate}

\hypertarget{uninstall}{%
\subsection{Uninstall}\label{uninstall}}

\begin{itemize}
\tightlist
\item
  uninstall \emph{MiKTeX}: delete from programs, then delete remaining
  folder in Program Files.
\item
  uninstall \emph{texlive}: go to \emph{c:/texlive}, delete the folder
\end{itemize}

\hypertarget{install}{%
\subsection{Install}\label{install}}

\emph{Install on Windows}

\begin{itemize}
\tightlist
\item
  \href{https://www.texstudio.org/}{Install Texstudio}
\item
  \href{https://www.tug.org/texlive/acquire-netinstall.html}{Install
  texlive}
\end{itemize}

\emph{Install on Ubuntu}

install texlive and texstudio ubuntu/debian.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{sudo}\NormalTok{ apt-get update}
\FunctionTok{sudo}\NormalTok{ apt-get install texlive-full}
\FunctionTok{sudo}\NormalTok{ apt-get install texstudio}
\end{Highlighting}
\end{Shaded}

\hypertarget{compile-in-atom-and-texstudio}{%
\subsection{Compile in Atom and
Texstudio}\label{compile-in-atom-and-texstudio}}

Install the following two packages in atom:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \href{https://github.com/thomasjo/atom-latex}{latex} package on atom.
\item
  \href{https://atom.io/packages/language-latex}{language-latex} package
  on atom
\end{enumerate}

In Atom Latex setting, note that on different platforms, paths to
texlive are different:

\begin{itemize}
\tightlist
\item
  Path:

  \begin{itemize}
  \tightlist
  \item
    Windows: \emph{C:/texlive/2019/bin/win32}
  \item
    Chromebook Linux: \emph{/usr/bin/latex}
  \end{itemize}
\item
  Which PDF viewer to use:

  \begin{itemize}
  \tightlist
  \item
    download Okular and use \href{https://okular.kde.org/}{Okular}
  \end{itemize}
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# set TeX path}
\CommentTok{# C:/texlive/2019/bin/win32}

\CommentTok{# Latex, after installation}
\CommentTok{# see where texlive appears under root folder directories}
\FunctionTok{sudo}\NormalTok{ find / -name }\StringTok{"texlive"}
\CommentTok{# see latex version, if texlive is used}
\ExtensionTok{latex}\NormalTok{ -v}
\CommentTok{# see main exe directory}
\FunctionTok{which}\NormalTok{ latex}
\CommentTok{# /usr/bin/latex}
\end{Highlighting}
\end{Shaded}

Go to Tex4Econ and compile various template files and other types of
files there to see if tex is working.

Compile the same files in texstudio.

\hypertarget{other-programs-to-install}{%
\section{Other Programs to Install}\label{other-programs-to-install}}

\hypertarget{package-managers}{%
\subsection{Package Managers}\label{package-managers}}

\begin{itemize}
\tightlist
\item
  \href{https://github.com/lukesampson/scoop/wiki/Quick-Start}{scoop}
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# in powershell}
\CommentTok{# enter cmd}
\CommentTok{# type powershell in cmd to enter powershell from cmd}
\ExtensionTok{powershell}
\CommentTok{# permission}
\ExtensionTok{set-executionpolicy}\NormalTok{ remotesigned -scope currentuser}
\CommentTok{# install}
\ExtensionTok{iwr}\NormalTok{ -useb get.scoop.sh }\KeywordTok{|} \ExtensionTok{iex}
\end{Highlighting}
\end{Shaded}

\hypertarget{editors}{%
\subsection{Editors}\label{editors}}

In addition to Atom, which is pretty slow, and harder to run on lower
spec computers.

\hypertarget{vscode}{%
\subsubsection{VSCode}\label{vscode}}

VSCode is much more light-weight than Atom, does not have hydrogen.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Debian Install}
\ExtensionTok{curl}\NormalTok{ https://packages.microsoft.com/keys/microsoft.asc }\KeywordTok{|} \ExtensionTok{gpg}\NormalTok{ --dearmor }\OperatorTok{>}\NormalTok{ microsoft.gpg}
\FunctionTok{sudo}\NormalTok{ install -o root -g root -m 644 microsoft.gpg /usr/share/keyrings/microsoft-archive-keyring.gpg}
\FunctionTok{sudo}\NormalTok{ sh -c }\StringTok{'echo "deb [arch=amd64 signed-by=/usr/share/keyrings/microsoft-archive-keyring.gpg] https://packages.microsoft.com/repos/vscode stable main" > /etc/apt/sources.list.d/vscode.list'}
\FunctionTok{sudo}\NormalTok{ apt-get install apt-transport-https}
\FunctionTok{sudo}\NormalTok{ apt-get update}
\FunctionTok{sudo}\NormalTok{ apt-get install code }\CommentTok{# or code-insiders}

\CommentTok{# Change Font (https://www.reddit.com/r/Crostini/comments/cesxr0/vscode_ui_and_fonts_too_small_heres_how_to_fix_it/)}
\CommentTok{# modify: /usr/share/applications/code.desktop}
\ExtensionTok{nvim}\NormalTok{ /usr/share/applications/code.desktop}
\CommentTok{#replace this line: Exec=/usr/share/code/code --unity-launch %F}
\VariableTok{Exec=}\NormalTok{sommelier }\ExtensionTok{-X}\NormalTok{ --scale=0.8 --dpi=160 /usr/share/code/code }\StringTok{"--unity-launch %F"}
\CommentTok{#replace this line: Exec=/usr/share/code/code --new-window %F}
\VariableTok{Exec=}\NormalTok{sommelier }\ExtensionTok{-X}\NormalTok{ --scale=0.8 --dpi=160 /usr/share/code/code }\StringTok{"--new-window %F"}

\CommentTok{# Open File}
\CommentTok{# low density mode to see cleary}
\ExtensionTok{code}
\end{Highlighting}
\end{Shaded}

Extensions to Download for VSCode:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  \href{https://marketplace.visualstudio.com/items?itemName=ms-python.python}{Microsoft
  Python Extension}
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{ext}\NormalTok{ install ms-python.python}
\end{Highlighting}
\end{Shaded}

\hypertarget{vscode-shortcuts}{%
\paragraph{VSCode Shortcuts}\label{vscode-shortcuts}}

\begin{itemize}
\tightlist
\item
  Open up Settings: \emph{Ctrl + ,}
\item
  Preview Markdown in VSCode: \emph{Ctrl + Shift + V}
\item
  Open up left pane: \emph{Ctrl + B}
\item
  Wrap Text: \emph{alt + Z}, switch between wrap or not to save screen
  space for example.
\end{itemize}

\hypertarget{vscode-benefits}{%
\paragraph{VSCode Benefits}\label{vscode-benefits}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  open a file, if have workspace where the file is in, automatically
  expand tree and show where file is on left pane.
\end{enumerate}

\hypertarget{vscode-and-r}{%
\paragraph{VSCode and R}\label{vscode-and-r}}

VSCode works pretty well it seems with Rmd files. Rmd files are useful
to work through algorithm, examples in preparation for writing usable
functions. Rstudio has been often very slow for me, and I generally feel
like can not stretch around and work in RStudio IDE.

In Atom, Hydrogen works, but really only works with very basic line by
line testing.

Using VSCode with RMD files seems much quicker, and there is room to
breath.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Install R following R installation instructions.
\item
  Then install extension
  \href{https://marketplace.visualstudio.com/items?itemName=Ikuyadeu.r}{R}.
  Microsoft stopped supporting rmd in VSCode:
  \href{https://github.com/microsoft/RTVS}{R Tools for Visual Studio}
\item
  properly set path (might have to call vscode from r env)

  \begin{itemize}
  \tightlist
  \item
    \emph{r.rterm.linux}:
    \emph{/home/wangfanbsg75/anaconda3/envs/r\_env/bin/R}
  \end{itemize}
\item
  Run:

  \begin{itemize}
  \tightlist
  \item
    ctrl + enter to activate R terminal
  \item
    VSCode 2019 allows for Knit Rmd: \emph{Ctrl + Shift + K}
  \item
    also can run code segment: select lines and \emph{ctrl + enter}
  \end{itemize}
\end{enumerate}

\hypertarget{vscode-and-latex}{%
\paragraph{VSCode and Latex}\label{vscode-and-latex}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  download
  \href{https://marketplace.visualstudio.com/items?itemName=James-Yu.latex-workshop}{Latex
  Workshop}
\end{enumerate}

\hypertarget{notepad}{%
\subsubsection{Notepad++}\label{notepad}}

\href{https://notepad-plus-plus.org/downloads/v7.8.2/}{Notepad++}. Many
old projects have xml folder structures for Notepad++.

\begin{itemize}
\tightlist
\item
  Silently update file changes made in other editors: Settings,
  Preferences, MISC., File Status Auto Detection, Update Silently.
\end{itemize}

\hypertarget{sublime}{%
\subsubsection{Sublime}\label{sublime}}

\href{https://www.sublimetext.com/3}{Sublime}

there are some issues with atom. When a file is edited inside Vim, when
changes are saved, a
\href{https://github.com/atom/atom/issues/17186}{temp file is created in
atom}. Sublime would, however, still show the file with the changes.
This makes dual screen editing very difficult. Sublime text is used for
this situation.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Sublime Linux}
\FunctionTok{wget}\NormalTok{ -qO - https://download.sublimetext.com/sublimehq-pub.gpg }\KeywordTok{|} \FunctionTok{sudo}\NormalTok{ apt-key add -}
\FunctionTok{sudo}\NormalTok{ apt-get install apt-transport-https}
\BuiltInTok{echo} \StringTok{"deb https://download.sublimetext.com/ apt/stable/"} \KeywordTok{|} \FunctionTok{sudo}\NormalTok{ tee /etc/apt/sources.list.d/sublime-text.list}
\FunctionTok{sudo}\NormalTok{ apt-get update}
\FunctionTok{sudo}\NormalTok{ apt-get install sublime-text}
\CommentTok{# launch from terminal}
\ExtensionTok{subl}
\end{Highlighting}
\end{Shaded}

\emph{Sublime Setup}:

\begin{itemize}
\tightlist
\item
  \href{https://weibeld.net/r/rmd_sublime_package.html}{r and rmd setup}
\end{itemize}

\hypertarget{pycharm}{%
\subsubsection{PyCharm}\label{pycharm}}

\href{https://www.}{pycharm}jetbrains.com/pycharm/)

\hypertarget{key-programs}{%
\subsection{Key Programs}\label{key-programs}}

\textbf{PDF}

\begin{itemize}
\tightlist
\item
  PDF Okular

  \begin{itemize}
  \tightlist
  \item
    Linux: \emph{sudo apt-get install okular}
  \item
    Windows: install okular from windows app store
  \item
    Okular if menu bar lost: \emph{Ctrl + m}
  \end{itemize}
\item
  PDF Adobe Acrobat

  \begin{itemize}
  \tightlist
  \item
    go to adobe website, log in using
    \href{mailto:fwang26@uh.edu}{\nolinkurl{fwang26@uh.edu}}
  \item
    go to my account, choose view and download my apps, choose acrobat
    \& PDF, download Acrobat DC
  \item
    UH account can only be activated on two accounts at once, so need to
    kick other computers out temporarily potentially
  \end{itemize}
\end{itemize}

\textbf{VPN}

\begin{itemize}
\tightlist
\item
  \href{https://protonvpn.com/}{Proton VPN}: unlimited usage free
  version available.
\item
  Express VPN
\end{itemize}

\hypertarget{utilities}{%
\subsection{Utilities}\label{utilities}}

\begin{itemize}
\tightlist
\item
  7-zip
\item
  evernote
\item
  keepass
\end{itemize}

\end{document}