### 2020-01-03 nontex/install/git/fs_git_setup.tex
set up compile minor update

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Git and Github Example and Tutorial, Sync with Multiple Repositories, Store Password},
            colorlinks=true,
            linkcolor=Maroon,
            citecolor=Blue,
            urlcolor=blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{Git and Github Example and Tutorial, Sync with Multiple Repositories,
Store Password}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

Go back to \href{http://fanwangecon.github.io/}{fan}'s
\href{http://fanwangecon.github.io/Tex4Econ/}{Tex4Econ and Miscellaneous
Repository}.

\hypertarget{download-programs}{%
\section{Download Programs}\label{download-programs}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Install \href{https://git-scm.com/download/win}{git for windows}:
  after install, try ``cd \textasciitilde{}/PyFan''
\item
  Install \href{https://atom.io/}{atom}
\item
  Might need to install \href{https://www.putty.org/}{putty} possibly
\end{enumerate}

\hypertarget{git-and-github-security-set-up}{%
\section{git and Github Security
Set-up}\label{git-and-github-security-set-up}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  open up git bash
\item
  generate rsa

  \begin{itemize}
  \tightlist
  \item
    \emph{ssh-keygen -t rsa -C
    ``\href{mailto:wangfanbsg75@live.com}{\nolinkurl{wangfanbsg75@live.com}}''}
  \item
    when prompted, do not enter ``file in which save the key'', when
    prompted for passphrase, enter ``WHATEVERPASSWORDIS''
  \end{itemize}
\item
  copy key

  \begin{itemize}
  \tightlist
  \item
    WINDOWS: \emph{clip \textless{} \textasciitilde{}/.ssh/id\_rsa.pub}
  \item
    LINUX: \emph{cat \textasciitilde{}/.ssh/id\_rsa.pub}
  \end{itemize}
\item
  log on to github ssh section, generate new ssh rsa key

  \begin{itemize}
  \tightlist
  \item
    SSH and GPG keys, choose New SSH key, paste in clipped text.
  \end{itemize}
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# inside git bash}
\FunctionTok{ssh-keygen}\NormalTok{ -t rsa -C }\StringTok{"wangfanbsg75@live.com"}
\CommentTok{# this copies the text in the .pub file generated}
\ExtensionTok{clip} \OperatorTok{<}\NormalTok{ ~/.ssh/id_rsa.pub}
\end{Highlighting}
\end{Shaded}

\hypertarget{start-and-sync-key-repositories}{%
\section{Start and Sync Key
Repositories}\label{start-and-sync-key-repositories}}

Inside git bash (open as \emph{administrator}), Sync Several Key
Repositories that should be synced on all computers:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  fanwangecon.github.io: Github root repo
\item
  PyFan: Private Repo with Python Support Files
\item
  \href{https://github.com/FanWangEcon/Tex4Econ}{Tex4Econ}: Latex,
  installation, and various other support files, public repo.
\item
  \href{https://github.com/FanWangEcon/R4Econ}{R4Econ}: Public repo, R
  package, R research programs
\item
  \href{https://github.com/FanWangEcon/M4Econ}{M4Econ}: Public repo,
  Matlab package, Matlab examples and research files
\item
  \href{https://github.com/FanWangEcon/Py4Econ}{Py4Econ}: Public repo,
  Python package, Python research programs
\item
  Teaching: Private teaching repository
\end{enumerate}

Other repositories can be synced when needed on an ad-hoc basis. The
repositories above are essential repositories.

\hypertarget{repo-folders-and-global-config}{%
\subsection{Repo Folders and Global
config}\label{repo-folders-and-global-config}}

Remember git bash or bash should be in windows opened as
\emph{administrator}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# cd to root folder}
\BuiltInTok{cd}\NormalTok{ ~}
\CommentTok{# generate all needed key repositories}
\FunctionTok{mkdir}\NormalTok{ fanwangecon.github.io PyFan Tex4Econ R4Econ M4Econ Py4Econ Teaching}

\CommentTok{# Set global config settings, in ~/.gitconfig}
\FunctionTok{git}\NormalTok{ config --global user.name }\StringTok{"Fan Wang"}
\FunctionTok{git}\NormalTok{ config --global user.email wangfanbsg75@live.com}
\end{Highlighting}
\end{Shaded}

\hypertarget{initalize-and-add-remote-repo-for-each}{%
\subsection{Initalize and add remote repo for
each:}\label{initalize-and-add-remote-repo-for-each}}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Initialize Repositories}
\BuiltInTok{cd}\NormalTok{ ~/fanwangecon.github.io}
\FunctionTok{git}\NormalTok{ init}
\FunctionTok{git}\NormalTok{ remote add github git@github.com:fanwangecon/fanwangecon.github.io.git}
\BuiltInTok{cd}\NormalTok{ ~/PyFan}
\FunctionTok{git}\NormalTok{ init}
\FunctionTok{git}\NormalTok{ remote add github git@github.com:fanwangecon/PyFan}
\BuiltInTok{cd}\NormalTok{ ~/Tex4Econ}
\FunctionTok{git}\NormalTok{ init}
\FunctionTok{git}\NormalTok{ remote add github git@github.com:fanwangecon/Tex4Econ}
\BuiltInTok{cd}\NormalTok{ ~/R4Econ}
\FunctionTok{git}\NormalTok{ init}
\FunctionTok{git}\NormalTok{ remote add github git@github.com:fanwangecon/R4Econ}
\BuiltInTok{cd}\NormalTok{ ~/M4Econ}
\FunctionTok{git}\NormalTok{ init}
\FunctionTok{git}\NormalTok{ remote add github git@github.com:fanwangecon/M4Econ}
\BuiltInTok{cd}\NormalTok{ ~/Py4Econ}
\FunctionTok{git}\NormalTok{ init}
\FunctionTok{git}\NormalTok{ remote add github git@github.com:fanwangecon/Py4Econ}
\BuiltInTok{cd}\NormalTok{ ~/Teaching}
\FunctionTok{git}\NormalTok{ init}
\FunctionTok{git}\NormalTok{ remote add github git@github.com:fanwangecon/Teaching}
\end{Highlighting}
\end{Shaded}

\hypertarget{sync-repo-in-ssh-secure-session}{%
\subsection{Sync Repo in SSH Secure
Session}\label{sync-repo-in-ssh-secure-session}}

Pull latested from multiple repositories. Just Paste the following
lines.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  start ssh-agent secure session
\item
  pull from multiple Repositories upon start of working session.
\item
  commit changes as work on one computer
\item
  upon leaving a computer with commited changes, push all
\item
  create bash file to make it easier to repeat steps
\end{enumerate}

Start and end ssh session:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# To avoid having to enter password each time, start background authentication agent.}
\BuiltInTok{eval} \StringTok{"}\VariableTok{$(}\FunctionTok{ssh-agent}\VariableTok{)}\StringTok{"}
\FunctionTok{ssh-add}\NormalTok{ ~/.ssh/id_rsa}

\CommentTok{# to stop the ssh session}
\BuiltInTok{kill} \VariableTok{$SSH_AGENT_PID}
\end{Highlighting}
\end{Shaded}

Pull from repos:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Pull from Repositories, do these one by one first}
\BuiltInTok{cd}\NormalTok{ ~/fanwangecon.github.io}
\FunctionTok{git}\NormalTok{ pull github master}
\BuiltInTok{cd}\NormalTok{ ~/PyFan}
\FunctionTok{git}\NormalTok{ pull github master}
\BuiltInTok{cd}\NormalTok{ ~/Tex4Econ}
\FunctionTok{git}\NormalTok{ pull github master}
\BuiltInTok{cd}\NormalTok{ ~/R4Econ}
\FunctionTok{git}\NormalTok{ pull github master}
\BuiltInTok{cd}\NormalTok{ ~/M4Econ}
\FunctionTok{git}\NormalTok{ pull github master}
\BuiltInTok{cd}\NormalTok{ ~/Py4Econ}
\FunctionTok{git}\NormalTok{ pull github master}
\BuiltInTok{cd}\NormalTok{ ~/Teaching}
\FunctionTok{git}\NormalTok{ pull github master}
\end{Highlighting}
\end{Shaded}

Push from repos committed changes

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Push to Repositories, do these one by one first}
\BuiltInTok{cd}\NormalTok{ ~/fanwangecon.github.io}
\FunctionTok{git}\NormalTok{ push -u github master}
\BuiltInTok{cd}\NormalTok{ ~/PyFan}
\FunctionTok{git}\NormalTok{ push -u github master}
\BuiltInTok{cd}\NormalTok{ ~/Tex4Econ}
\FunctionTok{git}\NormalTok{ push -u github master}
\BuiltInTok{cd}\NormalTok{ ~/R4Econ}
\FunctionTok{git}\NormalTok{ push -u github master}
\BuiltInTok{cd}\NormalTok{ ~/M4Econ}
\FunctionTok{git}\NormalTok{ push -u github master}
\BuiltInTok{cd}\NormalTok{ ~/Py4Econ}
\FunctionTok{git}\NormalTok{ push -u github master}
\BuiltInTok{cd}\NormalTok{ ~/Teaching}
\FunctionTok{git}\NormalTok{ push -u github master}
\end{Highlighting}
\end{Shaded}

to create bash file to store these (from git bash in windows):

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# store bash file in ~\textbackslash{}PyFan\textbackslash{}bin}
\FunctionTok{mkdir}\NormalTok{ ~/PyFan/bin}
\BuiltInTok{cd}\NormalTok{ ~/PyFan/bin}

\CommentTok{# create bash file}
\ExtensionTok{vim}\NormalTok{ pull_repos}
\ExtensionTok{vim}\NormalTok{ push_repos}

\CommentTok{# paste the text below over, comfirm bash loc with: which bash}
\CommentTok{#!/usr/bin/bash}
\BuiltInTok{echo}\NormalTok{ start fan github pull}
\CommentTok{# Security}
\BuiltInTok{eval} \StringTok{"}\VariableTok{$(}\FunctionTok{ssh-agent}\VariableTok{)}\StringTok{"}
\FunctionTok{ssh-add}\NormalTok{ ~/.ssh/id_rsa}
\CommentTok{# Pull from Persistent Repos}
\BuiltInTok{cd}\NormalTok{ ~/fanwangecon.github.io}
\FunctionTok{git}\NormalTok{ pull github master}
\CommentTok{# Stop Secured SSH}
\BuiltInTok{kill} \VariableTok{$SSH_AGENT_PID}

\CommentTok{# change permission to make file an executable}
\FunctionTok{chmod}\NormalTok{ u+x ~/PyFan/bin/pull_repos}
\FunctionTok{chmod}\NormalTok{ u+x ~/PyFan/bin/push_repos}
\CommentTok{# execute script}
\FunctionTok{bash}\NormalTok{ ~/PyFan/bin/pull_repos}
\end{Highlighting}
\end{Shaded}

\end{document}