### 2020-01-05 nontex/rmd/fex_rmd_template_guide.Rmd

---
title: |
  | Template Example Generation  
  | Example to do ABC
output:
  pdf_document:
    df_print: paged
    number_sections: true
  html_notebook:
    number_sections: true
  word_document:
    number_sections: true
  html_document:
    number_sections: true
urlcolor: blue
always_allow_html: yes
---

Go back to [fan](http://fanwangecon.github.io/)'s [Tex4Econ and Miscellaneous Repository](http://fanwangecon.github.io/Tex4Econ/).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Objective

Install various compilers and IDE:

1. Objective One
2. Objectie Two

# Part One

Use conda across platforms, so that locally on windows and ubuntu and remotely on aws, can have the same software setup environment.

- Search for Anaconda Prompt, right click, choose run as administrator.
- Check software versions.

```{bash}
conda list anaconda
python -V
```