### 2020-01-18 nontex/revealjs/_guide/fs_revealjs_basics.Rmd
revealjs markdown test

---
title: |
  | Basics of Reveal.js
  | for Mathmatical Presentations
output:
  pdf_document:
    df_print: paged
    number_sections: true
  html_notebook:
    number_sections: true
  word_document:
    number_sections: true
  html_document:
    number_sections: true
urlcolor: blue
always_allow_html: yes
---

Go back to [fan](http://fanwangecon.github.io/)'s [Tex4Econ and Miscellaneous Repository](http://fanwangecon.github.io/Tex4Econ/).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Equations and Reveal.js

## Page Fragments

- long html multiple files
- html fragments separate files
- rmd bookdown revealjs