### 2020-01-27 nontex/windows/tricks/fs_windows_tricks.Rmd
Equation gap control specific group

---
title: |
  | Windows Tricks
  | Shortcuts
output:
  pdf_document:
    df_print: paged
    number_sections: true
  html_notebook:
    number_sections: true
  word_document:
    number_sections: true
  html_document:
    number_sections: true
urlcolor: blue
always_allow_html: yes
---

Go back to [fan](http://fanwangecon.github.io/)'s [Tex4Econ and Miscellaneous Repository](http://fanwangecon.github.io/Tex4Econ/).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Windows Shortcuts

[Microsoft Keyboard shortcuts](https://www.microsoft.com/accessories/en-us/support/hardware-tips-and-tricks)

## Launch Programs Pinned to Task Bar

Can pin 1 through 0 programs to windows task bar, and access quickly via **win + 1** for the first pinned program for example.

But how to access the 11th pinned program?

**Info**:

- [5 Ways To Quickly Launch Programs On Windows](https://www.howtogeek.com/124109/5-ways-to-quickly-launch-programs-on-windows/)
- [Hey, Scripting Guy! How Can I Add a Windows PowerShell Shortcut to the Quick Launch Toolbar?](https://devblogs.microsoft.com/scripting/hey-scripting-guy-how-can-i-add-a-windows-powershell-shortcut-to-the-quick-launch-toolbar/)
- [Define shortcut key for shortcut in PowerShell](https://stackoverflow.com/questions/32126697/define-shortcut-key-for-shortcut-in-powershell)
- [get target path of shortcuts](https://social.technet.microsoft.com/Forums/en-US/fddb6ec0-cc48-43f0-929e-bf25d07fbf48/get-target-path-of-shortcuts?forum=winserverpowershell)