### 2020-02-04 nontex/revealjsrmd/test_img.Rmd
revealjs slides rmd with many images and to pdf with decktape

---
title: "Test Include Several Images in Revealjs and Generate PDF"
author: Fan Wang (fanwangecon.github.io)
date: February 4th, 2020
output:
  revealjs::revealjs_presentation:
    theme: sky
    highlight: pygments
    center: false
    transition: none
    css: fan_present.css
    mathjax: local
    self_contained: false
params:
    imgwidth: 25px
---

# Rmd RevealJS and Decktape
<!--
decktape convert to PDF
cd "C:/Users/fan/Tex4Econ/nontex/revealjsrmd/"
decktape automatic test_img.html test_img.pdf
-->

To produce very low cost, easy to compile, easy to produce presentations, use these:

1. [revealjs](https://revealjs.com/#/)
2. [revealjs_presentation](https://github.com/rstudio/revealjs)
3. [decktape](https://github.com/astefanutti/decktape)

Use the revealjs r package to write revealjs documents in r markdown and knit to html. Then use decktape to conver html to pdf.

## revealjs and r

See the [rmd file](http://fanwangecon.github.io/Tex4Econ/nontex/revealjsrmd/test_img.Rmd) that generated this presentation.

## revealjs and r

<!-- cd "C:/Users/fan/Tex4Econ/nontex/revealjsrmd/" -->

Once the HTML file is created, we can use decktape to convert it to pdf:

- install decktape: install [nodejs](https://nodejs.org/en/), npm install -g decktape
- to save results as image slides individually:
    * *decktape reveal --screenshots test_img.html test_img.pdf*
- to save results as a single pdf
    * *decktape reveal --size='2048x1536' test_img.html test_img.pdf*
    * [gap issue with large size workaround](https://github.com/astefanutti/decktape/issues/151)

# Image Group One

## Image One JPG

- small jpg profile photo
- Equation: $1+2=3$

![](images/profile_tiny.jpg)

## Two JPGs Together

- Two Images Together Same Page
- Equation: $1+2=3$

![](images/profile_tiny.jpg){width=100px}
![](images/github_logo.jpg){width=100px}

# Image Group Two

## pDF and SVG

- svg and then pdf

![](images/github_logo.svg){width=100px}
![Alt](images/github_logo.pdf){width=100px}