### 2020-02-06 nontex/revealjsrmd/test_tex.Rmd
revealjs newcommand tex math

---
title: "rmd revealjs Latex test"
author: Fan Wang 
date: Feb 5, 2020
output:
  revealjs::revealjs_presentation:
    theme: sky
    highlight: pygments
    center: true
    transition: none
    css: fan_present.css
    mathjax: local
    self_contained: false
---

# Overview

See more from [Fan](fanwangecon.github.io)'s [Tex4Econ](fanwangecon.github.io/Tex4Econ)

We will test out writing equations in RMD + revealjs

# Defining NEWCOMMAND

\newcommand{\vara}{\mathrm{Var}}
\newcommand{\varb}{\mathrm{\alpha + \beta}}
\newcommand{\varc}{
  \frac{a + b}{c + d} \times \exp\left( x \right) = y
}

<section style="font-size:60%;">

- This is from *\\vara+2*: $\vara+2$ 
- This is from *\\varb+2*: $\varb+2$ 
- This is from *\\varc+2*: $\varc+2$

</section>

```{tex echo=TRUE, eval=FALSE}
\newcommand{\vara}{\mathrm{Var}}
\newcommand{\varb}{\mathrm{\alpha + \beta}}
\newcommand{\varc}{
  \frac{a + b}{c + d} \times \exp\left( x \right) = y
}
```

# Equations

## Inline Equation

<div style="color:darkred;margin:20px;margin-top:50px;" >
Here is some text that is in red, 
<b>in between the b symbols mean put this text in bold</b> 
but this text is not bold
</div>

<p style="margin:10px;font-size:50%;">
<i>This is smaller italisized text</i>, font size 50 percent.
</p>

- Regular sized Equation: $1+2=3$
- Smaller Equation: <span style="font-size:50%;"> $1+2=3$ </span>

## Display Equation

<div style="font-size:40%;" >
$$
\begin{equation}
\label{eq:targetcost}
Z\left(\tau,\delta\right) =
\sum\limits_{
	\substack{
	\mathrm{cohort} \\ \in{\left\{70,72,74,76\right\}}}
	}
\left\{\delta\cdot
\int_{\epsilon}
\int_{Y_{min}}^{F_{Y}^{-1}\left(\tau\right)}
\int_{X}
N\Big(
\substack{
	Y,X,\epsilon; \\
	\delta, \Gamma_{\mathrm{cohort}}
}
\Big)f\left(X|Y\right)f\left(Y\right)f\left(\epsilon\right)\mathrm{d}X\mathrm{d}Y\mathrm{d}\epsilon\right\}
\end{equation}
$$
</div>

# Equations New Command

\newcommand{\target}{
\begin{equation}
Z\left(\tau,\delta\right) =
\sum\limits_{
	\substack{
	\mathrm{cohort} \\ \in{\left\{70,72,74,76\right\}}}
	}
\left\{\delta\cdot
\int_{\epsilon}
\int_{Y_{min}}^{F_{Y}^{-1}\left(\tau\right)}
\int_{X}
N\Big(
\substack{
	Y,X,\epsilon; \\
	\delta, \Gamma_{\mathrm{cohort}}
}
\Big)f\left(X|Y\right)f\left(Y\right)f\left(\epsilon\right)\mathrm{d}X\mathrm{d}Y\mathrm{d}\epsilon\right\}
\end{equation}
}

Equation defined as new command with different zoom:

<span style="font-size:10%;"> $$\target$$ </span>
<span style="font-size:20%;"> $$\target$$ </span>
<span style="font-size:30%;"> $$\target$$ </span>
<span style="font-size:40%;"> $$\target$$ </span>
<span style="font-size:50%;"> $$\target$$ </span>