### 2020-04-20 nontex/git/g_basics/git_basics.Rmd
python poetry installation guides

---
title: "Basic Git Commands"
output:
  pdf_document: default
  word_document: default
  html_document: default
  html_notebook: default
urlcolor: blue
always_allow_html: yes
---

### Basic Git Commands

#### Check File Status

Check if a particular file has had local changes that have not been commited yet.

Use the [*git status*](https://git-scm.com/docs/git-status) command, with short *-s* option:

- *modified*: the *grepl(' M ')* check will be true
- *new*: the *grepl('?? ')* check will be true
- *existing no change*: the *(st_git_status == "")* check will be true

```{r}
spg_modified <- 
  paste0('git status -s',
         'C:/Users/fan/R4Econ/amto/array/fs_ary_mesh.Rmd')
spg_anewfile <- 
  paste0('git status -s',
         'C:/Users/fan/R4Econ/math/differentiation/fs_differentiate_lhopital.Rmd')
spg_nochange <- 
  paste0('git status -s',
         'C:/Users/fan/R4Econ/math/func_ineq/fs_atkinson_ces.Rmd')

st_git_status <- toString(system(spg_modified, intern=TRUE))
grepl(' M ', st_git_status, fixed=TRUE)
grepl('?? ', st_git_status, fixed=TRUE)
(st_git_status == "")

st_git_status <- toString(system(spg_anewfile, intern=TRUE))
grepl(' M ', st_git_status, fixed=TRUE)
grepl('?? ', st_git_status, fixed=TRUE)
(st_git_status == "")

st_git_status <- toString(system(spg_nochange, intern=TRUE))
grepl(' M ', st_git_status, fixed=TRUE)
grepl('?? ', st_git_status, fixed=TRUE)
(st_git_status == "")
```