### 2020-04-28 _support/docu/hierarchy/flatten.Rmd
over and unber bar, bold or not. bm mand accents package.

---
title: "Tex Flatten Subdirectory files to Root"
titleshort: "Tex Flatten Subdirectory files to Root"
description: |
  Working directory has multiple subfolders, folders, and refer to input files, flatten all into one level without subfolders.
  Extract subdirectory files to main directory and edit main file links.
core:
  - package: tex
    code: |
      tex
date: 2020-04-27
date_start: 2020-04-27
output:
  pdf_document:
    pandoc_args: '../../_output_kniti_pdf.yaml'
    includes:
      in_header: '../../preamble.tex'
  html_document:
    pandoc_args: '../../_output_kniti_html.yaml'
    includes:
      in_header: "../../hdga.html"
always_allow_html: true
urlcolor: blue
---

### Flatten Latex File

```{r global_options, include = FALSE}
try(source("../../.Rprofile"))
```

`r text_shared_preamble_one`
`r text_shared_preamble_two`
`r text_shared_preamble_thr`

- latex automatically extract \input to put in same file
- [Replace \input{fileX} by the content of fileX automatically](https://tex.stackexchange.com/questions/21838/replace-inputfilex-by-the-content-of-filex-automatically/21840)

Can use Pandoc, but pandoc conversion leads to all incorrect labels, confusing references.

```{bash}
cd "C:\Users\fan\Documents\Dropbox (UH-ECON)\Project Emily Minority Survey\Drafts_RR\ChinaSchoolClosure"
pandoc -s draft_edcc_submit_rr.tex -o draft_edcc_submit_rr_flat.tex
```

use latexpand

```{bash}
conda activate wk_perl
cd "C:\Users\fan\Documents\Dropbox (UH-ECON)\Project Emily Minority Survey\Drafts_RR\OverleafChinaSchoolClosure"
perl "C:\Users\fan\.conda\envs\wk_perl\latexpand\latexpand" draft_main.tex > draft_main_flat.tex
```