### 2020-05-15 _support/docu/fs_comment.tex
timelime improvements, bold symbol, etc. revealjs tests.

\documentclass[12pt,english]{article}
\usepackage[utf8]{inputenc}

% url package
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,
    urlcolor=cyan,
}
% Bibliography package
\usepackage[authordate, backend=biber]{biblatex-chicago}
\bibliography{fansample}

% Titling and Author
\title{Latex Preamble}
\author{\href{https://fanwangecon.github.io/}{Fan Wang}\thanks{https://fanwangecon.github.io, repository: \href{https://fanwangecon.github.io/Tex4Econ/}{Tex4Econ}}}
\date{\today}

\begin{document}

Comment will not appear with percent in front

\begin{verbatim}
% Comment
\end{verbatim}
% Comment

yml block?

\begin{verbatim}
<!-- regular html comment -->
\end{verbatim}

<!-- regular html comment -->

\end{document}