### 2020-05-15 nontex/revealjsrmd/math_tests/fragments/preamble_main.tex

\input{../fragments/preamble_one.tex}
\input{../fragments/preamble_two.tex}

% Special String Shorthands
\input{../fragments/shorthand/short_text_syms.tex}
\input{../fragments/shorthand/short_math.tex}
\input{../fragments/shorthand/short_math_param.tex}
\input{../fragments/shorthand/short_math_var.tex}
\input{../fragments/shorthand/short_math_slopeintercept.tex}
\input{../fragments/shorthand/short_math_esti_space.tex}
\input{../fragments/shorthand/short_math_ces_prod.tex}

\input{../fragments/shorthand/short_math_ntw.tex}
\input{../fragments/shorthand/short_support.tex}