### 2020-05-15 nontex/revealjsrmd/math_tests/fragments/shorthand/short_support.tex

% Various supporting strings
% \newcommand{\papertitle}{Optimal Targeting Queues}
% \newcommand{\papertitle}{From Efficiency to Equity: the Optimal Allocation of Discrete and Continuous Resources Among Heterogeneous Individuals}
\newcommand{\papertitle}{The Optimal Allocation of Resources Among Heterogeneous Individuals}
% \newcommand{\papertitle}{The Allocative Implications of Empirical Estimates}
\newcommand{\paperauthor}{\href{https://fanwangecon.github.io}{Fan Wang}}
\newcommand{\urlpaper}{https://fanwangecon.github.io/assets/FanWang_OptimalTargetingQueue.pdf}
\newcommand{\urlrepo}{https://fanwangecon.github.io/PrjOptiAlloc/}
\newcommand{\thankszpart}[1]{This is a file that contains all #1\xspace from the paper ``\papertitle''. The paper's supporting website is \href{\urlrepo}{https://fanwangecon.github.io/PrjOptiAlloc/}. The full paper pdf can be found \href{https://fanwangecon.github.io/assets/FanWang_OptimalTargetingQueues.pdf}{here}}
\newcommand{\thanksmain}{Department of Economics, University of Houston (email: fwang26@uh.edu). The latest paper pdf can be found \href{https://fanwangecon.github.io/assets/FanWang_OptimalTargetingQueue.pdf}{here}. The paper's supporting website is \href{\urlrepo}{https://fanwangecon.github.io/PrjOptiAlloc/}.}