### 2020-11-22 nontex/development/python/fn_installations.Rmd
additional installation guides

---
title: |
  | Develop Python Package with Sphinx
output:
  html_document:
    number_sections: true
    toc: yes
  html_notebook:
    number_sections: true
    toc: yes
  word_document:
    number_sections: true
  pdf_document:
    df_print: paged
    number_sections: true
    toc: yes
urlcolor: blue
always_allow_html: yes
---

Go back to [fan](http://fanwangecon.github.io/)'s [Tex4Econ and Miscellaneous Repository](http://fanwangecon.github.io/Tex4Econ/).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Develop Python Package with Sphinx

## RST

[cheatsheet](https://github.com/ralsina/rst-cheatsheet/blob/master/rst-cheatsheet.rst)

### Sections

```
Chapter 1 Title
===============
Section 1.1 Title
-----------------
Subsection 1.1.1 Title
~~~~~~~~~~~~~~~~~~~~~~
Section 1.2 Title
-----------------
Chapter 2 Title
===============
```

### Definitions

*literal block*: ?