### 2020-11-22 nontex/install/windows/fn_installations_R.Rmd

---
title: |
  | R Installation and Update
output:
  html_document:
    number_sections: true
    toc: yes
  html_notebook:
    number_sections: true
    toc: yes
  word_document:
    number_sections: true
  pdf_document:
    df_print: paged
    number_sections: true
    toc: yes
urlcolor: blue
always_allow_html: yes
---

Go back to [fan](http://fanwangecon.github.io/)'s [Tex4Econ and Miscellaneous Repository](http://fanwangecon.github.io/Tex4Econ/).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# R Installation and Set-Up

Install R and Update R

## Fully Uninstall r

Uninstall R, RStudio and RTools from Windows "Programs and Features" menu

If R was installed in a virtual environment, delete the environment. Otherwise, use system's uninstaller. Before that, from terminal/command-prompt, type *R*, and *.libPaths()* to find paths. Do so inside conda main, outside of conda main, inside different environments, find all paths. After uninstaller finishes, check in the libPath folders to see if there are still stuff there, delete all, delete the folders.

```{r}
ls_spn_paths <- .libPaths()
print(ls_spn_paths)
# C:/Users/fan/Documents/R/win-library/3.6
# C:/Program Files/R/R-3.6.1/library
```

## Install R First Time

1. [download R](https://cloud.r-project.org/)
    + for debian: [Johannes Ranke](https://cran.r-project.org/bin/linux/debian/). For Linux/Debian installation, crucial to update the *source.list* to include sources that have more recent versions of R. If not, will get very old R versions that is not compatible with many packages.
    + add R to path for Windows. In Windows Path, add for example: *C:/Program Files/R/R-3.6.2/bin/x64/* and *C:/Rtools/bin*
2. [Install Rtools](https://cran.r-project.org/bin/windows/Rtools/) for building R packages.
2. [download R-studio](https://rstudio.com/products/rstudio/download/)
3. Open R-studio and auto-detect R
4. Install additional packages

After installing R, change the path sequence so that packages install for all users:

```{r}
ls_spn_paths <- .libPaths()
print(ls_spn_paths)
ls_spn_paths <- c(ls_spn_paths[2], ls_spn_paths[1])
.libPaths(ls_spn_paths)
ls_spn_paths <- .libPaths()
print(ls_spn_paths)
```

```{r}
# Install RTools First!
# https://cran.r-project.org/bin/windows/Rtools/

# Install system tools
install.packages(c("backports"))

# Install tidyverse
install.packages(c("tidyverse", "tidymodels", "vroom"))

# Install Packaging tools
install.packages(c("devtools", "pkgdown", "roxygen2", "bookdown", "knitr", "kableExtra", "formatR", "revealjs"))

# Install Statistics models
install.packages(c("AER", "minpack.lm"))
install.packages(c("quantreg"))

# Install Tools to Work with Other Packages
install.packages(c("reticulate", "JuliaCall"))
# for reticulate errors, install directly from: devtools::install_github("rstudio/reticulate")

# Paralell
install.packages(c("parallel", "doParallel", "foreach"))

# for sublime r-ide
# install.packages("languageserver")

# Personal Packages
devtools::install_github("fanwangecon/REconTools")
devtools::install_github("fanwangecon/PrjOptiAlloc")

# Stata in Rmd
devtools::install_github("Hemken/Statamarkdown")
```

Temp Installs:
```
# 2020-10-19
# Temp install development version due to but
# https://github.com/rstudio/reticulate/issues/831
devtools::install_github("rstudio/reticulate")
```

## Update R on Computer

1. On windows, install the *installr* package, and use *updateR()*
2. At the end, will ask if want to move all old pacages to new R directory

New R will have new package directory, could keep all in old, should copy all to new, and not keep old.

```{r}
# https://www.r-project.org/nosvn/pandoc/installr.html
install.packages('installr')
# update R from inside R (not Rstudio)
require(installr)
# this will open dialog boxes to take you through the steps.
updateR()
# Set Rstudio to the Latest R
```

## R Tests

Test the following file to see if we can execute a R file. Do it inside *r_env* and inside a *r* session.

```{r}
# # A simple file with summary statistics using tidyverse
# source('C:/Users/fan/R4Econ/summarize/dist/htmlpdfr/fst_hist_onevar.R')
# # Another simple file with summary statistics using tidyverse
# source('C:/Users/fan/R4Econ/support/tibble/htmlpdfr/fs_tib_basics.R')
# # A file involving estimation
# source('C:/Users/fan/R4Econ/optimization/cesloglin/htmlpdfr/fst_ces_plan_linlog.R')
#
# C:/Users/fan/R4Econ/summarize/dist/fst_hist_onevar.Rmd
# C:/Users/fan/R4Econ/support/tibble/fs_tib_basics.Rmd
# C:/Users/fan/R4Econ/optimization/cesloglin/fst_ces_plan_linlog.Rmd
```