### 2022-03-09 template/fantemplate_acceptedsubmit/manuscript_tab_fig.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figures
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Figure 1
\newcommand{\subFigWidth}{0.50}
\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{\subFigWidth\textwidth}
        \centering
        \includegraphics[width=\linewidth]{\string"fig_main/fig1a_Tester_Indiff_person_495_med_2021".eps}
        \caption{Contours of indifference curves at given estimates}
    \end{subfigure}~
	\begin{subfigure}[b]{\subFigWidth\textwidth}
	    \centering
    	\includegraphics[width=\linewidth]{\string"fig_main/fig1b_Tester_Indiff_person_495_lambdam3_20210130".eps}
    	\caption{Vary $\lambda$}
	\end{subfigure}
	\par\medskip
	\begin{subfigure}[b]{\subFigWidth\textwidth}
        \centering
        \includegraphics[width=\linewidth]{\string"fig_main/fig1c_Tester_Indiff_person_495_refmeanm3_20210130".eps}
        \caption{Vary $\mu_{R}$}
    \end{subfigure}~
	\begin{subfigure}[b]{\subFigWidth\textwidth}
	    \centering
	    \includegraphics[width=\linewidth]{\string"fig_main/fig1d_Tester_Indiff_person_495_refsd_increasem3_20210130".eps}
	    \caption{Vary $\sigma_{R}$}
	\end{subfigure}
	\caption{Consumption and height outcome indifference curves and choice frontiers. \emph{Note}: Health possibility frontier and indifference curves are visualized in blue (solid-line) given estimated parameters and characteristics for one particular child. The consumption and expected height outcome frontier is determined by the budget constraint as well as the production function. Panels b, c and d show results from varying one parameter at a time in red (dashed-line) and orange (dotted-line). See Section \ref{sec:maximize} for discussions.}
	\label{fig:indiff}
\end{figure}
\pagebreak

% Figure 2
\begin{figure}[H]
    \centering
    \begin{subfigure}[t]{.43\textwidth}
        \centering
        \includegraphics[width=\linewidth]{\string"fig_main/fig2a_ss_mean_hgt24_yr_trimmed".eps}
        \caption{Height}
    \end{subfigure}~
	\begin{subfigure}[t]{0.57\textwidth}
	    \centering
    	\includegraphics[width=\linewidth]{\string"fig_main/fig2b_ss_mean_prot_15t24_yr".eps}
    	\caption{Protein}
	\end{subfigure}
	\caption{Increasing protein and height gaps across cohorts. \emph{Note}: See Section \ref{sec:gap} for discussions.}
	\label{fig:PortHgtGap}
\end{figure}
\pagebreak

\clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Tables
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Table 1
\input{"tab_main/tab1_summ_stats_covar_main_keep.tex"}
\clearpage
% Table 2
\input{tab_main/tab2_estimates_2018}
\clearpage
% Table 3
\input{"tab_main/tab3_modelfit.tex"}
\clearpage
% Table 4
\input{"tab_main/tab4_frescoatoledecompose.tex"}
\clearpage
% Table 5
\input{tab_main/tab5_estimates_2021}
\clearpage
\pagebreak