### 2022-03-09 template/fantemplate_acceptedsubmit/tab_main/tab4_frescoatoledecompose.tex

\begin{table}[htbp]
\centering                                 \def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}                                 \caption{Decompose the protein gap between Atole and Fresco villages\label{frescoatoledecompose}}                                 \begin{adjustbox}{max width=1\textwidth}                                 \begin{tabular}{m{2.5cm} >{\centering\arraybackslash}m{1.8cm} >{\centering\arraybackslash}m{1.8cm} >{\centering\arraybackslash}m{1.8cm} >{\centering\arraybackslash}m{1.8cm} >{\centering\arraybackslash}m{1.8cm}}                                 \toprule
& \multicolumn{5}{C{9cm}}{Average protein choice across birth years} \\
\cmidrule(l{5pt}r{5pt}){2-6}
& \multicolumn{1}{C{1.8cm}}{\small \textbf{Fresco}} & \multicolumn{3}{C{5.4cm}}{\small \textbf{Fresco counterfactuals}} & \multicolumn{1}{C{1.8cm}}{\small \textbf{Atole}} \\
\cmidrule(l{5pt}r{5pt}){2-2} \cmidrule(l{5pt}r{5pt}){3-5} \cmidrule(l{5pt}r{5pt}){6-6}
& \multicolumn{1}{C{1.8cm}}{\textit{\small simulated without counterfactuals}} & \multicolumn{1}{C{1.8cm}}{\textit{\small Fresco with Atole price discount}} & \multicolumn{1}{C{1.8cm}}{\textit{\small Fresco with Atole ref. point}} & \multicolumn{1}{C{1.8cm}}{\textit{\small Fresco with both Atole price and ref. point}} & \multicolumn{1}{C{1.8cm}}{\textit{\small simulated without counterfactuals}} \\                                         
\midrule
\multicolumn{6}{L{13.3cm}}{\vspace*{-5mm}\hspace*{-5mm}\textbf{\normalsize Panel a: Height across birth cohorts}} \\
&            &            &            &            &            \\
1970-71             &       76.52&       77.20&       77.09&       77.68&       77.71\\
1972-73             &       76.80&       77.35&       77.57&       78.08&       78.07\\
1974-75             &       76.89&       77.48&       77.86&       78.43&       78.65\\
% \midrule
% Observations        &        7050&        7050&        7050&        7050&        8070\\
\midrule
\multicolumn{6}{L{13.3cm}}{\vspace*{-5mm}\hspace*{-5mm}\textbf{\normalsize Panel b: Protein across birth cohorts}} \\                                          &            &            &            &            &            \\
1970-71             &       18.43&       20.80&       20.48&       22.69&       23.13\\
1972-73             &       19.30&       21.35&       22.27&       24.33&       24.51\\
1974-75             &       19.97&       22.13&       23.95&       26.32&       27.13\\
% \midrule
% Observations        &        7050&        7050&        7050&        7050&        8070\\
\bottomrule
\addlinespace[0.5em]
\multicolumn{6}{l}{\footnotesize{\emph{Note}: See Section \ref{sec:decompose} for discussions.}}\\
\addlinespace
\end{tabular}
\end{adjustbox}
\end{table}