### 2022-04-21 nontex/git/g_basics/fs_git_bash.Rmd
git basics and git bash pull single repo

---
title: "BASH Pull and Push From GIT Remote Repository"
output:
  pdf_document: default
  word_document: default
  html_document: default
  html_notebook: default
urlcolor: blue
always_allow_html: yes
---

Go back to [Fan](http://fanwangecon.github.io/)'s [Tex4Econ and Miscellaneous Repository](http://fanwangecon.github.io/Tex4Econ/).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

### Basic BASH

#### Pull from Remote Repo and Create a New Local Directory

[Sample pull bash file](https://fanwangecon.github.io/Tex4Econ/nontex/git/g_basics/fs_git_bash_pull_one) is shown below.

```{}
#!/usr/bin/bash

# This bash file, after making the appropriate adjustments pulls from a remote
# project and creates a new local folder.

# ADJUT 1 STRREPO
# STRREPO='Py4Econ'
STRREPO='PrjHealthProductionCGW'

if [ -d "/g/repos/" ]
then
  echo "Computer has g drive, on precision"
  STRREPOWTHPATH="/g/repos/${STRREPO}"
else
  echo "Computer has c drive, on vostro"
  STRREPOWTHPATH="/c/Users/fan/${STRREPO}"
fi
echo "${STRREPOWTHPATH}"

echo pull single repository
echo going to pull from $STRREPOWTHPATH

# generate all needed key repositories
mkdir $STRREPOWTHPATH

# Initialize
cd $STRREPOWTHPATH

# ADJUT 2.1 ACCOUNT
# Set global config settings, in ~/.gitconfig
git config --global user.name "Fan Wang"
git config --global user.email wangfanbsg75@live.com

# Initialize
git init
# ADJUT 2.2 ACCOUNT
git remote add github git@github.com:fanwangecon/$STRREPO

# Secure
eval "$(ssh-agent)"
# ssh-add ~/.ssh/id_rsa_yy
ssh-add ~/.ssh/id_rsa

# Pull
cd $STRREPOWTHPATH
# ADJUT 3 root name
git pull github main
# git pull github master

# Open Repository in Atom
atom $STRREPOWTHPATH
```