### 2025-11-09 _other/tikz/tree/fs_tikz_nodediagram.tex

\documentclass{article}

\usepackage{blindtext}
\usepackage{multicol}
\usepackage{caption}
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{decorations.text}

\usepackage{pgfplots}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,
    urlcolor=cyan,
}
\usepackage{geometry}
\geometry{
	a4paper,
	noheadfoot=true,
	left=1.0in,
	right=1.0in,
	top=1.0in,
	bottom=1.0in
}

% url package
\usepackage{hyperref}
\usepackage{subcaption}
\usepackage{relsize} % make some math large
\usepackage{bm} % make some math bold

% Itemize formating
\usepackage{enumitem}

% Figure positioning
\usepackage{float}

% Titling and Author
\title{Latex Tikz Examples, Node-based Diagrams}
\author{\href{https://fanwangecon.github.io/}{Fan Wang}\thanks{\url{https://fanwangecon.github.io}, repository: \href{https://fanwangecon.github.io/Tex4Econ/}{Tex4Econ}}}
\date{\today}

\begin{document}

\maketitle

\section{Connecting Nodes with Arrows}

In the following example, there are three sets of data, relevant for three research objectives:

\begin{verbatim}
    \begin{tikzpicture}[thick]
        \node(1) at (-1,-1) [rectangle,draw=blue!50,fill=blue!20] {Data 1};
        \node(2) at (7 ,-2) [rectangle,draw=blue!50,fill=blue!20] {Data 2};
        \node(3) at (-1,-4) [rectangle,draw=blue!50,fill=blue!20] {Data 3};

        \node(4) at (3,-1) [rectangle,draw=black!50] {Analysis Objective 1};
        \node(5) at (3,-2) [rectangle,draw=black!50] {Analysis Objective 2};
        \node(6) at (3,-3) [rectangle,draw=black!50] {Analysis Objective 3};

        \draw [->] (1) edge (4);
        \draw [->] (1) |- (5);
        \draw [->] (1) |- (6);

        \draw [->] (2) edge (5);
        \draw [->] (2) |- (6);

        \draw [->] (3) -| (6);
    \end{tikzpicture}    
\end{verbatim}

\medskip
\begin{center}
    \begin{tikzpicture}[thick]
        \node(1) at (-1,-1) [rectangle,draw=blue!50,fill=blue!20] {Data 1};
        \node(2) at (7 ,-2) [rectangle,draw=blue!50,fill=blue!20] {Data 2};
        \node(3) at (-1,-4) [rectangle,draw=blue!50,fill=blue!20] {Data 3};

        \node(4) at (3,-1) [rectangle,draw=black!50] {Analysis Objective 1};
        \node(5) at (3,-2) [rectangle,draw=black!50] {Analysis Objective 2};
        \node(6) at (3,-3) [rectangle,draw=black!50] {Analysis Objective 3};

        \draw [->] (1) edge (4);
        \draw [->] (1) |- (5);
        \draw [->] (1) |- (6);

        \draw [->] (2) edge (5);
        \draw [->] (2) |- (6);

        \draw [->] (3) -| (6);
    \end{tikzpicture}
\end{center}
\clearpage

\section{Connecting Blocks of Nodes and with Blocks of Nodes}

To present the relationship between data and research analysis, show visually how several research aim will be achieved with different data structures and analysis.

\begin{verbatim}
    \begin{tikzpicture}[thick, font=\small]
        \node(100) at (2.5,10) [rectangle] {Research Aim I};
        \node(101) at (-3,8.5) [rectangle, draw=blue!50, fill=blue!20] {Data 1};
        \node(102) at (-3,7.5) [rectangle, draw=black!50, fill=red!20] {Data 2};
        \node(111) at (0,8) [rectangle, draw=black!50] {Combine};
        \draw [->] (101) edge (111);
        \draw [->] (102) edge (111);
        \node(121) at (3,9) [rectangle, draw=black!50] {Analysis I};
        \node(122) at (3,8) [rectangle, draw=black!50] {Analysis II};
        \node(123) at (3,7) [rectangle, draw=black!50] {Analysis III};
        \draw [->] (111) |- (121);
        \draw [->] (111) |- (122);
        \draw [->] (111) |- (123);
        \node(131) at (8,8) [rectangle, draw=black!50, text width=5cm] {Some Text.};
    \end{tikzpicture}    
\end{verbatim}

\begin{center}
\begin{tikzpicture}[thick, font=\small]
    % Objective I
    \node(100) at (2.5,10) [rectangle] {Research Aim I};
    \node(101) at (-3,8.5) [rectangle, draw=blue!50, fill=blue!20] {Data 1};
    \node(102) at (-3,7.5) [rectangle, draw=black!50, fill=red!20] {Data 2};

    \node(111) at (0,8) [rectangle, draw=black!50] {Combine};
    \draw [->] (101) edge (111);
    \draw [->] (102) edge (111);

    \node(121) at (3,9) [rectangle, draw=black!50] {Analysis I};
    \node(122) at (3,8) [rectangle, draw=black!50] {Analysis II};
    \node(123) at (3,7) [rectangle, draw=black!50] {Analysis III};
    \draw [->] (111) |- (121);
    \draw [->] (111) |- (122);
    \draw [->] (111) |- (123);

    \node(131) at (8,8) [rectangle, draw=black!50, text width=5cm] {specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.};

    % Objective II
    \node(200) at (2.5,5) [rectangle] {Research Aim II};
    \node(201) at (-3,3.5) [rectangle, draw=blue!50, fill=blue!20] {Data 1};
    \node(202) at (-3,2.5) [rectangle, draw=black!50, fill=green!20] {Data 3};

    \node(211) at (0,3) [rectangle, draw=black!50] {Combine};
    \draw [->] (201) edge (211);
    \draw [->] (202) edge (211);

    \node(221) at (3,4) [rectangle, draw=black!50] {Analysis I};
    \node(222) at (3,3) [rectangle, draw=black!50] {Analysis II};
    \node(223) at (3,2) [rectangle, draw=black!50] {Analysis III};
    \draw [->] (211) |- (221);
    \draw [->] (211) |- (222);
    \draw [->] (211) |- (223);

    \node(231) at (8,3) [rectangle, draw=black!50, text width=5cm] {specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.};

    % Objective III
    \node(300) at (2.5,0) [rectangle] {Research Aim III};
    \node(301) at (-3,-1.5) [rectangle, draw=blue!50, fill=blue!20] {Data 1};
    \node(302) at (-3,-2.5) [rectangle, draw=black!50, fill=brown!20] {Data 4};

    \node(311) at (0,-2) [rectangle, draw=black!50] {Combine};
    \draw [->] (301) edge (311);
    \draw [->] (302) edge (311);

    \node(321) at (3,-1.5) [rectangle, draw=black!50] {Analysis I};
    \node(322) at (3,-2.5) [rectangle, draw=black!50] {Analysis II};
    \draw [->] (311) |- (321);
    \draw [->] (311) |- (322);

    \node(331) at (8,-2) [rectangle, draw=black!50, text width=5cm] {specimen book. It has survived not only five centuries, but also the leap into electronic typesetting, remaining essentially unchanged.};

\end{tikzpicture}
\end{center}
\clearpage
\newpage

\section{Concentric Circle Diagrams}

\subsection{Single Set of Concentric Circles}

We have three circles concentric, with word on top. Here is the code:
\begin{verbatim}
\begin{tikzpicture}[thick, font=\small]
\def\sizechd{1cm}
\def\sizehum{3cm}
\def\sizefac{4.5cm}
    % Objective I
    \node(101) at (0,0) [fill, circle, 
        draw=black!50, fill=yellow!40, 
        minimum size=\sizefac] {};
    \node(102) at (0,0) [fill, circle, 
        draw=blue!50, fill=blue!20, 
        minimum size=\sizehum, label=above:Home] {};
    \node(103) at (0,0) [circle, 
        draw=black!50, fill=white!20, 
        scale=1, minimum size=\sizechd, label=above:Caretakers] {Children};
\end{tikzpicture}
\end{verbatim}

Here is the figure:
\begin{center}
\begin{tikzpicture}[thick, font=\small]
\def\sizechd{1cm}
\def\sizehum{3cm}
\def\sizefac{4.5cm}
    % Objective I
    \node(101) at (0,0) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {Data 3};
    \node(102) at (0,0) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum, label=above:Home] {Data 2};
    \node(103) at (0,0) [circle, draw=black!50, fill=white!20, scale=1, minimum size=\sizechd, label=above:Caretakers] {Children};
\end{tikzpicture}
\end{center}

\subsection{Three Sets of Concentric Circles}

We have three sets of three circles concentric, with word on top.
\begin{center}
\begin{tikzpicture}[thick, font=\small]
\def\sizechd{1cm}
\def\sizehum{3cm}
\def\sizefac{4.5cm}
\def\xcira{-2.15}
\def\xcirb{2.15}
\def\xcirc{0}
\def\ycira{0}
\def\ycirb{0}
\def\ycirc{-3.75}

% Circle A
\node(101) at (\xcira, \ycira) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {};
\node(102) at (\xcira, \ycira) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum, label=above:Home] {};
\node(103) at (\xcira, \ycira) [circle, draw=black!50, fill=white!20, scale=1, minimum size=\sizechd, label=above:Caretakers] {Children};
% \path   [thin, draw=blue, text=black,    % <--- new
%          postaction={decorate},           % <--- changed
%          text effects={text along path}]  % <--- new
%         (4,0) arc [start angle=-60,end angle=80,radius=4];

% Circle B
\node(101) at (\xcirb, \ycirb) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {Data 3};
\node(102) at (\xcirb, \ycirb) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum, label=above:Home] {Data 2};
\node(103) at (\xcirb, \ycirb) [circle, draw=black!50, fill=white!20, scale=1, minimum size=\sizechd, label=above:Caretakers] {Children};

% Circle C
\node(101) at (\xcirc, \ycirc) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {Data 3};
\node(102) at (\xcirc, \ycirc) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum, label=above:Home] {Data 2};
\node(103) at (\xcirc, \ycirc) [circle, draw=black!50, fill=white!20, scale=1, minimum size=\sizechd, label=above:Caretakers] {Children};
\end{tikzpicture}
\end{center}

\subsection{Circle with Circular Text}

We have three circles concentric, with concentric text. The text starts from 180 degree and goes to 0 degree, covering half a circle. The text is centered, so if there is less text, the text appears in the center.

\begin{center}
\begin{tikzpicture}[thick, font=\small]
\def\sizechd{1cm}
\def\sizehum{3cm}
\def\sizefac{4.5cm}
\def\xcira{0}
\def\ycira{0}

% Circle A
\node(101) at (\xcira, \ycira) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {};
\node(102) at (\xcira, \ycira) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum] {};
\node(103) at (\xcira, \ycira) [circle, draw=black!50, fill=white!20, scale=1, minimum size=\sizechd] {Children};

\draw[decoration={
    text along path,
    text={Caretakers are here.},
    text align={center},
    raise=0cm},decorate] (-1, 0) arc (180:0:1);
\draw[decoration={
    text along path,
    text={This is our home!},
    text align={center},
    raise=0cm},decorate] (-1.8, 0) arc (180:0:1.8);

\end{tikzpicture}
\end{center}

\subsection{Three Sets of Concentric Circles with Circular Text}

We have three sets of three circles concentric, with word on top.

\begin{center}
\begin{tikzpicture}[thick, font=\normalsize]
\def\sizechd{1.5cm}
\def\sizehum{3.5cm}
\def\sizefac{5cm}
% Cirlce x and y
\def\xcira{-2.4}
\def\xcirb{2.4}
\def\xcirc{0}
\def\ycira{0}
\def\ycirb{0}
\def\ycirc{-4.2}

% Circle A
\node(101) at (\xcira, \ycira) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {};
\node(102) at (\xcira, \ycira) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum] {};
\node(103) at (\xcira, \ycira) [circle, draw=black!50, fill=white!20, scale=1, minimum size=\sizechd] {Children};
\draw[decoration={
    text along path,
    text={Child health professionals},
    text align={center},
    raise=0cm},decorate] (-3.60, 0) arc (180:0:1.22);
\draw[decoration={
    text along path,
    text={Health facilities},
    text align={center},
    raise=0cm},decorate] (-4.4, 0) arc (180:0:2);

% Circle B
\node(101) at (\xcirb, \ycirb) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {Data 3};
\node(102) at (\xcirb, \ycirb) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum] {Data 2};
\node(103) at (\xcirb, \ycirb) [circle, draw=black!50, fill=white!20, scale=1, minimum size=\sizechd] {Children};
\draw[decoration={
    text along path,
    text={Teachers Teachers Teachers Teachers Teachers.},
    text align={center},
    raise=0cm},decorate] (1.20, 0) arc (180:-179:1.22);
\draw[decoration={
    text along path,
    text={Schools Schools Schools Schools Schools Schools Schools Schools Schools.},
    text align={center},
    raise=0cm},decorate] (0.4, 0) arc (180:-179:2);

% Circle C
\node(101) at (\xcirc, \ycirc) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {Data 3};
\node(102) at (\xcirc, \ycirc) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum] {Data 2};
\node(103) at (\xcirc, \ycirc) [circle, draw=black!50, fill=white!20, scale=1, minimum size=\sizechd] {Children};
\draw[decoration={
    text along path,
    text={Teachers Teachers Teachers Teachers Teachers.},
    text align={center},
    raise=0cm},decorate] (-1.22, -4.20) arc (180:-179:1.22);
\draw[decoration={
    text along path,
    text={Schools Schools Schools Schools Schools Schools Schools Schools Schools.},
    text align={center},
    raise=0cm},decorate] (-2.02, -4.20) arc (180:-179:2);

\end{tikzpicture}
\end{center}
\clearpage

\subsection{Three Sets of Concentric Circles with Circular Text and Big Eclipse}

We have three sets of three circles concentric, with word on top.

\begin{center}
\begin{tikzpicture}[thick, font=\normalsize]
\def\sizechd{1.5cm}
\def\sizehum{3.5cm}
\def\sizefac{5cm}
% Cirlce x and y
\def\xcira{-2.4}
\def\xcirb{2.4}
\def\xcirc{0}
\def\ycira{0}
\def\ycirb{0}
\def\ycirc{-4.2}

\draw [black, fill=gray!20] (0, -1.7) ellipse (6.75cm and 5.5cm);

% Circle A
\node(101) at (\xcira, \ycira) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {};
\node(102) at (\xcira, \ycira) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum] {};
\node(103) at (\xcira, \ycira) [circle, draw=black!50, fill=white!20, scale=1, minimum size=\sizechd] {Children};
\draw[decoration={
    text along path,
    text={Child health professionals},
    text align={center},
    raise=0cm},decorate] (-3.60, 0) arc (180:0:1.22);
\draw[decoration={
    text along path,
    text={Health facilities},
    text align={center},
    raise=0cm},decorate] (-4.4, 0) arc (180:0:2);

% Circle B
\node(101) at (\xcirb, \ycirb) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {Data 3};
\node(102) at (\xcirb, \ycirb) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum] {Data 2};
\node(103) at (\xcirb, \ycirb) [circle, draw=black!50, fill=white!20, scale=1, minimum size=\sizechd] {Children};
\draw[decoration={
    text along path,
    text={Teachers Teachers Teachers Teachers Teachers.},
    text align={center},
    raise=0cm},decorate] (1.20, 0) arc (180:-179:1.22);
\draw[decoration={
    text along path,
    text={Schools Schools Schools Schools Schools Schools Schools Schools Schools.},
    text align={center},
    raise=0cm},decorate] (0.4, 0) arc (180:-179:2);

% Circle C
\node(101) at (\xcirc, \ycirc) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {Data 3};
\node(102) at (\xcirc, \ycirc) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum] {Data 2};
\node(103) at (\xcirc, \ycirc) [circle, draw=black!50, fill=white!20, scale=1, minimum size=\sizechd] {Children};
\draw[decoration={
    text along path,
    text={Teachers Teachers Teachers Teachers Teachers.},
    text align={center},
    raise=0cm},decorate] (-1.22, -4.20) arc (180:-179:1.22);
\draw[decoration={
    text along path,
    text={Schools Schools Schools Schools Schools Schools Schools Schools Schools.},
    text align={center},
    raise=0cm},decorate] (-2.02, -4.20) arc (180:-179:2);

% Text 
% \node[below] at (0, 3.5) {
% Climatic environment, extreme weather, etc.};
\node[below] at (0, 3.5) [rectangle, draw=gray!50, fill=gray!10, text width=3cm] {Climate shocks environmental crisis.};

\end{tikzpicture}
\end{center}
\clearpage

\subsection{Three Sets of Concentric Circles with Circular Text and Big Eclipse and Arrows Outcomes}

We have three sets of three circles concentric, with word on top. Edges and centered text below.

\begin{center}
\begin{tikzpicture}[thick, font=\small]
\tikzstyle{every node}=[draw,shape=ellipse];
\def\sizechd{1.5cm}
\def\sizehum{3.5cm}
\def\sizefac{5cm}
% Cirlce x and y
\def\xcira{-2.4}
\def\xcirb{2.4}
\def\xcirc{0}
\def\ycira{0}
\def\ycirb{0}
\def\ycirc{-4.2}

% Background gray
\draw [black, fill=gray!20] (0, -1.4) ellipse (5.75cm and 5.75cm);

% Circle A
\node(101) at (\xcira, \ycira) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {};
\node(102) at (\xcira, \ycira) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum] {};
\node(103) at (\xcira, \ycira) [circle, draw=black!50, fill=white!20, scale=1, minimum size=\sizechd] {Children};
\draw[decoration={
    text along path,
    text={Child health professionals},
    text align={center},
    raise=0cm},decorate] (-3.60, 0) arc (180:0:1.22);
\draw[decoration={
    text along path,
    text={Health facilities},
    text align={center},
    raise=0cm},decorate] (-4.4, 0) arc (180:0:2);

% Circle B
\node(201) at (\xcirb, \ycirb) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {};
\node(202) at (\xcirb, \ycirb) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum] {};
\node(203) at (\xcirb, \ycirb) [circle, draw=black!50, fill=white!20, scale=1, minimum size=\sizechd] {Children};
\draw[decoration={
    text along path,
    text={Teachers Teachers Teachers Teachers Teachers.},
    text align={center},
    raise=0cm},decorate] (1.20, 0) arc (180:-179:1.22);
\draw[decoration={
    text along path,
    text={Schools Schools Schools Schools Schools Schools Schools Schools Schools.},
    text align={center},
    raise=0cm},decorate] (0.4, 0) arc (180:-179:2);

% Circle C
\node(301) at (\xcirc, \ycirc) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {};
\node(302) at (\xcirc, \ycirc) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum] {};
\node(303) at (\xcirc, \ycirc) [circle, draw=black!50, fill=white!20, scale=1, minimum size=\sizechd] {Children};
\draw[decoration={
    text along path,
    text={Teachers Teachers Teachers Teachers Teachers.},
    text align={center},
    raise=0cm},decorate] (-1.22, -4.20) arc (180:-179:1.22);
\draw[decoration={
    text along path,
    text={Schools Schools Schools Schools Schools Schools Schools Schools Schools.},
    text align={center},
    raise=0cm},decorate] (-2.02, -4.20) arc (180:-179:2);

% Text 
% \node[below] at (0, 3.5) {
% Climatic environment, extreme weather, etc.};
% \node[below] at (0, 3.5) [rectangle, draw=gray!50, fill=gray!10, text width=3.5cm] {\textbf{{\normalsize \textcolor{blue}{Extreme Weather Crisis}}}};
\draw[decoration={
    text along path,
    text={Extreme Weather Crisis Extreme Weather Crisis Extreme Weather Crisis Extreme Weather Crisis.},
    text align={center},
    raise=0cm},decorate] (-6, -1.45) arc (180:0:6);

\node(898) at (-4, -3.75) [rectangle, draw=gray!20, fill=gray!20, text width=0cm] {};
\node(899) at (+4, -3.75) [rectangle, draw=gray!20, fill=gray!20, text width=0cm] {};

\node at (0, 3.5) [rectangle, draw=blue!20, fill=gray!20, minimum width=3.8cm] {\textcolor{blue}{Extreme temperature}};
\node at (0, 2.9) [rectangle, draw=blue!20, fill=gray!20, minimum width=3.8cm] {\textcolor{blue}{Pollution}};
\node(998) at (0, -7.75) [rectangle, draw=gray!10, fill=gray!0, minimum width=3.8cm] {\textcolor{red}{Human Capital Formation}};
\node(999) at (0, -8.25) [rectangle, draw=gray!10, fill=gray!0, minimum width=7.1cm] {\textcolor{black}{Learning, psychological, socio-emotional, physical outcomes}};

\path[->] (101.south west) edge [double, out=-1200, 
    in=100, draw=red,line width=1pt] (898.north);
\path[->] (301.west) edge [double, out=-1200, 
    in=0, draw=red,line width=1pt] (898.east);
\path[->] (898.south west) edge [double, out=-1200, 
    in=150, draw=red,line width=1pt] (998.west);

\path[->] (201.south east) edge [double, out=-70, 
    in=100, draw=red,line width=1pt] (899.north);
\path[->] (301.east) edge [double, out=0, 
    in=-200, draw=red,line width=1pt] (899.west);
\path[->] (899.south east) edge [double, out=0, 
    in=20, draw=red,line width=1pt] (998.east);

\end{tikzpicture}
\end{center}
% \clearpage

\subsection{Three Sets of Concentric Circles with Circular Text and Big Eclipse and Arrows Outcomes and Outside Arrows and Central Circle}

We have three sets of three circles concentric, with word on top. Edges and centered text below. Additional circle inside, and centered areas. We also resize the area the make it smaller
% VSPACE reduction needed after adding the outside dashed lines, which somehow added extra spaces
\begin{center}
\vspace*{-1cm}
\hspace*{-0.5cm}
\resizebox{8cm}{!}{
\begin{tikzpicture}[font=\small]
\tikzstyle{every node}=[draw,shape=ellipse];
\def\sizechd{1.5cm}
\def\sizehum{3.5cm}
\def\sizefac{5cm}
% Cirlce x and y
\def\xcira{-2.4}
\def\xcirb{2.4}
\def\xcirc{0}
\def\ycira{0}
\def\ycirb{0}
\def\ycirc{-4.2}

% Background gray
\draw [black, fill=gray!20] (0, -1.4) ellipse (5.75cm and 5.75cm);

% Circle A
\node(101) at (\xcira, \ycira) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {};
\node(102) at (\xcira, \ycira) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum] {};
\node(103) at (\xcira, \ycira) [circle, draw=black!50, fill=white!20, scale=1, minimum size=\sizechd] {Children};
\draw[decoration={
    text along path,
    text={Child health professionals},
    text align={center},
    raise=0cm},decorate] (-3.60, 0) arc (180:0:1.22);
\draw[decoration={
    text along path,
    text={Maternal/Child health facilities},
    text align={center},
    raise=0cm},decorate] (-4.4, 0) arc (180:0:2);

% Circle B
\node(201) at (\xcirb, \ycirb) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {};
\node(202) at (\xcirb, \ycirb) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum] {};
\node(203) at (\xcirb, \ycirb) [circle, draw=blue!50, fill=white!20, scale=1, minimum size=\sizechd] {Children};
\draw[decoration={
    text along path,
    text={Teachers and curriculum},
    text align={center},
    raise=0cm},decorate] (1.20, 0) arc (180:0:1.22);
\draw[decoration={
    text along path,
    text={Pre-school, primary/secondary schools},
    text align={center},
    raise=0cm},decorate] (0.4, 0) arc (180:0:2);

% Circle C
\node(301) at (\xcirc, \ycirc) [fill, circle, draw=black!50, fill=yellow!40, minimum size=\sizefac] {};
\node(302) at (\xcirc, \ycirc) [fill, circle, draw=blue!50, fill=blue!20, minimum size=\sizehum] {};
\node(303) at (\xcirc, \ycirc) [circle, draw=blue!50, fill=white!20, scale=1, minimum size=\sizechd] {Children};
\draw[decoration={
    text along path,
    text={Parents and caretakers},
    text align={center},
    raise=0cm},decorate] (-1.3, -4.20) arc (-180:0:1.3);
\draw[decoration={
    text along path,
    text={Living environment and housing},
    text align={center},
    raise=0cm},decorate] (-2.2, -4.20) arc (-180:0:2.2);

% Circle D
\path [draw=red!50, fill=red!30, fill opacity = 1.0, line width=1pt, even odd rule] (0,-1.42) circle (1.5) (0,-1.42) circle (0.8);
\node(403) at (0,-1.42) [circle, draw=red!50, fill=white!20, scale=1, minimum size=\sizechd] {Children};
\draw[decoration={
    text along path,
    text={Policies and community},
    text align={center},
    raise=0cm},decorate] (-1.30, -1.35) arc (-180:0:1.25);

% Text 
% \node[below] at (0, 3.5) {
% Climatic environment, extreme weather, etc.};
% \node[below] at (0, 3.5) [rectangle, draw=gray!50, fill=gray!10, text width=3.5cm] {\textbf{{\normalsize \textcolor{blue}{Extreme Weather Crisis}}}};
\draw[decoration={
    text along path,
    text={Climatic ambient environmental exposures during childhood that condition human capital growth and development},
    text align={center},
    raise=0cm},decorate] (-6, -1.45) arc (180:0:6);

\node(898) at (-4, -3.75) [rectangle, draw=gray!20, fill=gray!20, text width=0cm] {};
\node(899) at (+4, -3.75) [rectangle, draw=gray!20, fill=gray!20, text width=0cm] {};

\node(996) at (0, 3.5) [rectangle, draw=black!20, fill=gray!20, minimum width=3.8cm, font=\bfseries] {\textcolor{black}{Extreme temperature}};
\node(997) at (0, 3.0) [rectangle, draw=black!20, fill=gray!20, minimum width=3.8cm, font=\bfseries] {\textcolor{black}{Pollution}};
\node(998) at (0, -7.75) [rectangle, draw=gray!10, fill=gray!0, minimum width=3.8cm] {\textcolor{red}{Human Capital Formation}};
\node(999) at (0, -8.25) [rectangle, draw=gray!10, fill=gray!0, minimum width=7.1cm] {\textcolor{black}{Learning, psychological, socio-emotional, physical outcomes}};

% to and out
% https://tikz.dev/library-edges
% out = The angle at which the curve leaves the start coordinate. If the start coordinate is a node, the start coordinate is the point on the border of the node at the given ⟨angle⟩. The control point will, thus, lie at a certain distance in the direction ⟨angle⟩ from the start coordinate.
% in = The angle at which the curve reaches the target coordinate.
\path[->] (101.south west) edge [double, out=-105, 
    in=100, draw=blue,line width=1pt] (898.north);
\path[->] (301.west) edge [double, out=-105, 
    in=0, draw=blue,line width=1pt] (898.east);
\path[->] (898.south west) edge [double, out=-105, 
    in=150, draw=blue,line width=1pt] (998.west);

\path[->] (201.south east) edge [double, 
    out=-70, in=100, 
    draw=blue,line width=1pt] (899.north);
\path[->] (301.east) edge [double, 
    out=0, in=-200, 
    draw=blue,line width=1pt] (899.west);
\path[->] (899.south east) edge [double, 
    out=-75, in=30, 
    draw=blue,line width=1pt] (998.east);        
\path[dashed, ->] (999.north east) edge [
    out=30, in=35, looseness=1
    ] (996.north east);
\path[dashed, ->] (999.north west) edge [
    out=150, in=145, looseness=1
    ] (996.north west);

\path[<->] (103.east) edge [double, draw=black, line width=1pt] (403.north west);
\path[<->] (203.west) edge [double, draw=black, line width=1pt] (403.north east);
\path[<->] (303.north) edge [double, draw=black, line width=1pt] (403.south);

\end{tikzpicture}
}
\end{center}
Above we have the picture we are interested in including.
\clearpage

\section{Rows of Data and Analysis Connected Nodes}

\subsection{Research Plan Diagram Example One}

To present the relationship between data and research analysis, show visually how several research aim will be achieved with different data structures and analysis.

\begin{figure}[H]
\centering
\caption{Research Aim I---Inequality in Ambient Environmental Exposures}
\begin{tikzpicture}[thick, font=\small]
% A. Sizing for tikz
\def\fstdata{4.2cm}
\def\fstanalysis{5.2cm}
\def\fstunequal{3.0cm}
\def\fstnote{13.2cm}
% B.1 Text for tikz group 1
\newcommand{\dataone}{\textbf{Environmental Data} $Z_{l,y}$: ambient exposure in location $l$, period $y$.}
\newcommand{\datatwo}{\textbf{Population Data} $P_{i,l,t}$: size of demographic group $i$ in location $l$, period $y$.}
% B.2 Text for tikz group 2
\newcommand{\analysisone}{\emph{Analysis I}: Excess pollution burden in each (or combined) location $l$ and over time.}
\newcommand{\analysistwo}{\emph{Analysis II}: Excess pollution burden for demographic group $i$ over time.}
\newcommand{\analysisthree}{\emph{Analysis III}: Excess pollution burden for demographic group $i$ in each (or combined) location $l$ over time.}
% B.3 Text for tikz group 3
\newcommand{\inequalitymeasure}{Exposure inequality: 
\begin{enumerate}[leftmargin=0cm,itemindent=.3cm,labelwidth=\itemindent,labelsep=0.0cm, itemsep=-0.1cm, topsep=0cm, align=left]
    \item Overall
    \item Across groups
    \item Within group
\end{enumerate}}
% C Tikz Draw
    % Objective I
    % \node(100) at (2.5,10.75) [rectangle] {};
    \node(101) at (-3,8.8) [rectangle, draw=blue!50, fill=blue!20, text width=\fstdata] {\dataone};
    \node(102) at (-3,7.2) [rectangle, draw=black!50, fill=red!20, text width=\fstdata] {\datatwo};

    \node(111) at (0.5,8) [circle, draw=black!50] {Merge};
    \draw [->] (101) edge (111);
    \draw [->] (102) edge (111);

    \node(121) at (4.5,9.5) [rectangle, draw=black!50, text width=\fstanalysis] {\analysisone};
    \node(122) at (4.5,8) [rectangle, draw=black!50, text width=\fstanalysis] {\analysistwo};
    \node(123) at (4.5,6.5) [rectangle, draw=black!50, text width=\fstanalysis] {\analysisthree};
    \draw [->] (111.north) |- (121);
    \draw [->] (111.east) |- (122);
    \draw [->] (111.south) |- (123);

    \node(131) at (9.5,8) [rectangle, draw=black!50, text width=\fstunequal] {\inequalitymeasure};
    \draw [->] (121) to [bend left=45] (131);
    \draw [->] (122) to (131);
    \draw [->] (123) to [bend right=45] (131);

    \node(141) at (2.5,2.25) [rectangle, draw=black!50, text width=\fstnote, label={Inequality distributional percentile calculations}, align=center] {
Given $M$ locations and $N$ demographic groups, let $\mathcal{P}_{i,y}=\sum_{l=1}^{M}
P_{l,i,y}$ be group $i$'s population and $\mathcal{Z}_{i,y}$ be group $i$'s mean exposure. The quantile function $Q_{x}(\tau)$ with $\tau\in[0,1)$ is:\\
$\begin{aligned}   
\text{Overall: } &Q_{Z_{y}}(\tau) =
\max_{l\in\left\{1, \dots, M \right\}}
\left(
    \text{Z}_{l,y}
    \times
    \mathbf{1}\left\{
    \tau \ge
      \sum_{\tilde{l}=1}^{M} P_{\tilde{l},y}
      \times
      \mathbf{1}\left\{
        Z_{\tilde{l},y} \le Z_{l,y}
        \right\}
    \right\}
\right)\\
\text{Across: } &Q_{\mathcal{Z}_{y}}(\tau) =
\max_{i\in\left\{1, \dots, N \right\}}
\left(
    \mathcal{Z}_{i,y}
    \times
    \mathbf{1}\left\{
    \tau \ge
      \sum_{\tilde{i}=1}^{N} \mathcal{P}_{\tilde{i},y}
      \times
      \mathbf{1}\left\{
        \mathcal{Z}_{\tilde{i},y} \le \mathcal{Z}_{i,y}
        \right\}
    \right\}
\right)\\
\text{Within: } &Q_{Z_{i,y}}(\tau) =
\max_{l\in\left\{1, \dots, M \right\}}
\left(
    \text{Z}_{l,y}
    \times
    \mathbf{1}\left\{
    \tau \ge
      \sum_{\tilde{l}=1}^{M}
      \frac{P_{i, \tilde{l}, y}}{\mathcal{P}_{i,y}}
      \times
      \mathbf{1}\left\{
        Z_{\tilde{l},y} \le Z_{l,y}
        \right\}
    \right\}
\right)
% , \forall i\in\left\{1,\dots,N\right\}
\end{aligned}$    
    };
\draw [->] (131.east) to [bend left=45] (141.east);

\end{tikzpicture}
\end{figure}
\clearpage

\subsection{Research Plan Diagram Example Two}

\begin{figure}[H]
\centering
\caption{Research Aim II---Inequality in Ambient Environmental Exposures}
\begin{tikzpicture}[thick, font=\small]
% A. Sizing for tikz
\def\fstdata{4.2cm}
\def\fstanalysis{4.5cm}
\def\fstunequal{3.5cm}
\def\fstnote{16.2cm}
% B.1 Text for tikz group 1
\newcommand{\dataone}{\textbf{Environmental Data} $Z_{l,y}$: ambient exposure in location $l$ during period $y$.}
\newcommand{\datatwo}{\textbf{Human Capital Data} $Y_{j,i,l,y}$: human capital outcomes for individual $j$ of demographic group $i$ in location $l$ during period $y$.}
\newcommand{\datathree}{\textbf{SES Status Data} $X_{j}$: Parental/household SES and other attributes for $j$.}
% B.2 Text for tikz group 2
\newcommand{\analysisone}{\emph{Analysis I}: Explain average variations in $Y_{j,i,l,t}$ with variations in $Z_{l,y}$ across $l$ and $y$.}
\newcommand{\analysistwo}{\emph{Analysis II}: Explain quantiles of residual variations in $Y_{j,i,l,t}$ with variations in $Z_{l,y}$ across $l$ and $y$.}
\newcommand{\analysisthree}{\emph{Analysis III}: Explain variations in the quantiles of $Y_{j,i,l,t}$ with variations in $Z_{l,y}$ across $l$ and $y$.}
% B.3 Text for tikz group 3
\newcommand{\inequalitymeasure}{
\textbf{Allow for}: 
\begin{enumerate}[leftmargin=0cm,itemindent=.3cm,labelwidth=\itemindent,labelsep=0cm, itemsep=-0.05cm, topsep=0cm, align=left]
    \item Multiple $Z_{l,y}$ 
    \item $Z_{ly} \cdot X_{i}$ interaction
    % \item Instruments (e.g. thermal inversion)
\end{enumerate}
\medskip
\textbf{Control for}: 
\begin{enumerate}[leftmargin=0cm,itemindent=.3cm,labelwidth=\itemindent,labelsep=0cm, itemsep=-0.05cm, topsep=0cm, align=left]
    \item Location $\times$ time trends
    \item Invariant demo. group attributes
    \item Pre-determined parental attributes
\end{enumerate}}
% C Tikz Draw
    % Objective I
    % \node(100) at (2.5,10.75) [rectangle] {};
    \node(101) at (-3,10) [rectangle, draw=blue!50, fill=blue!20, text width=\fstdata] {\dataone};
    \node(102) at (-3,8) [rectangle, draw=black!50, fill=green!20, text width=\fstdata] {\datatwo};
    \node(103) at (-3,6) [rectangle, draw=black!50, fill=yellow!20, text width=\fstdata] {\datathree};

    \node(111) at (0.5,8) [circle, draw=black!50] {Merge};
    \draw [->] (101.east) to (111);
    \draw [->] (102) edge (111);
    \draw [->] (103.east) to (111);

    \node(121) at (4.5,9.75) [rectangle, draw=black!50, text width=\fstanalysis] {\analysisone};
    \node(122) at (4.5,8) [rectangle, draw=black!50, text width=\fstanalysis] {\analysistwo};
    \node(123) at (4.5,6.1) [rectangle, draw=black!50, text width=\fstanalysis] {\analysisthree};
    \draw [->] (111.north) |- (121);
    \draw [->] (111.east) |- (122);
    \draw [->] (111.south) |- (123);

    \node(131) at (9.5,8) [rectangle, draw=black!50, text width=\fstunequal] {\inequalitymeasure};
    \draw [->] (121) to [bend left=45] (131);
    \draw [->] (122) to (131);
    \draw [->] (123) to [bend right=45] (131);

    \node(141) at (3.00,2.8) [rectangle, draw=black!50, text width=\fstnote, label={Mean and quantile estimations}, align=left] {To illustrate, assume an additive model, and let $\widehat{Y}_{jily}= 
\alpha_0 \cdot \widehat{Z}_{ly}
+ \alpha_1 \cdot \widehat{Z}_{ly} \cdot X_{j}
+ \alpha_2 \cdot X_{j}
+ \theta_{i}
+ \phi_{l}
+ \psi_{ly}$, 
where $\widehat{Z}$ are SES dimensions of parental characteristics and $\theta_{i}
+ \phi_{l}
+ \psi_{ly}$ are demographic group, location, and location-specific time fixed effects.
Analysis I minimizes 
$\sum_{j,i,l,y} \left(Y_{jily} - \widehat{Y}_{jily}\right)^2$. 
Analysis II minimizes 
$\sum_{j,i,l,y} 
\left(\tau-
\mathbf{1}
\left\{ Y_{jily} - \widehat{Y}_{jily} \le 0\right\}
\right) \cdot \left(Y_{jily} - \widehat{Y}_{jily}\right)
$, where $\mathbf{1}\left\{\cdot\right\}$ denotes the indicator function.
Analysis III minimizes 
$\sum_{j,i,l,y} 
\left(\widehat{RIF}\left(\widehat{Q(\tau)}, Y_{jily}\right) - 
\widehat{Y}_{jily}\right)^2$, where $\widehat{RIF}$ is the quantile-specific recentered influence function.
};
\draw [->] (131.east) to [bend left=45] (141.east);

\end{tikzpicture}
\end{figure}

\clearpage

\subsection{Research Plan Diagram Example Three}

\begin{figure}[H]
\centering
\caption{Research Aim III---Inequality in Ambient Environmental Exposures}
\begin{tikzpicture}[thick, font=\small]
% A. Sizing for tikz
\def\fstdata{4.2cm}
\def\fstanalysis{4.75cm}
\def\fstunequal{3.90cm}
\def\fstnote{16.2cm}
% B.1 Text for tikz group 1
\newcommand{\dataone}{\textbf{Environmental Data} $Z_{l,y}$: Ambient exposure in location $l$ during period $y$.}
\newcommand{\datatwo}{\textbf{Home Inputs} $H_{j,i,l,y}$: Home/parental human capital inputs for child $j$ of demo-SES group $i$ in location $l$ during period $y$.}
\newcommand{\datathree}{\textbf{Community Inputs} $S_{l,y}$: Community inputs (e.g. schools, health clinics) in location $l$ during period $y$.}
% B.2 Text for tikz group 2
\newcommand{\analysisone}{\emph{Analysis I}: In period $y$, explain variations in $H_{j,i,l,t}$ and $S_{l,y}$ with variations in $Z_{l,y}$ across $l$.}
\newcommand{\analysistwo}{\emph{Analysis II}:
Explain variations in $H_{j,i,l,y}$ and $S_{l,y}$ with the path of variations in $\left\{Z_{l,\widehat{y}}\right\}_{\widehat{y}=y-\tau}^{y}$ from period $y-\tau$ to $y$.
% Evaluate the persistence of the association between $Y_{j,i,l,y}$ and $Z_{l,y-\tau}$ from $\tau$ periods ago, considering $Z_{l,y}$.
}
\newcommand{\analysisthree}{\emph{Analysis III}: Explain variations in $Y_{j,i,l,y}$ (Human capital outcomes) with variations in $Y_{j,i,l,y-1}$, $H_{j,i,l,t}$, $S_{l,y}$, and $Z_{l,y}$ across $j$ and $l$.}
% B.3 Text for tikz group 3
\newcommand{\inequalitymeasure}{
\textbf{Endogeneity}: 
\begin{enumerate}[leftmargin=0cm,itemindent=.3cm,labelwidth=\itemindent,labelsep=0cm, itemsep=-0.05cm, topsep=0cm, align=left]
    % \item Multiple $Z_{l,y}$ 
    % \item $Z_{ly} \cdot X_{i}$ interaction
    % \item Isolate unpredictable components of $Z_{l,y}$ 
    \item $\mathrm{cov}\left(Z_{l,y},\text{Inputs}\right)\neq 0$ 
    \item $\mathrm{P}\left(Z_{l,y}|Z_{l,y-\tau}\right)\neq \mathrm{P}\left(Z_{l,y}\right)$
    % \item Instruments (e.g. thermal inversion)
\end{enumerate}
\medskip
\textbf{Identification}: 
\begin{enumerate}[leftmargin=0cm,itemindent=.3cm,labelwidth=\itemindent,labelsep=0cm, itemsep=-0.05cm, topsep=0cm, align=left]
    \item Consider multiple $Z_{l,y}$ and inputs
    \item Consider current and lagged inputs
    \item Instruments (e.g. thermal inversion)
\end{enumerate}}
% C Tikz Draw
    % Objective I
    % \node(100) at (2.5,10.75) [rectangle] {};
    \node(101) at (-3,9.9) [rectangle, draw=blue!50, fill=blue!20, text width=\fstdata] {\dataone};
    \node(102) at (-3,8.0) [rectangle, draw=black!50, fill=brown!20, text width=\fstdata] {\datatwo};
    \node(103) at (-3,5.9) [rectangle, draw=black!50, fill=orange!20, text width=\fstdata] {\datathree};

    \node(111) at (0.5,8) [circle, draw=black!50] {Merge};
    \draw [->] (101.east) to (111);
    \draw [->] (102) edge (111);
    \draw [->] (103.east) to (111);

    \node(121) at (4.3,9.7) [rectangle, draw=black!50, text width=\fstanalysis] {\analysisone};
    \node(122) at (4.3,8.0) [rectangle, draw=black!50, text width=\fstanalysis] {\analysistwo};
    \node(123) at (4.3,5.9) [rectangle, draw=black!50, text width=\fstanalysis] {\analysisthree};
    \draw [->] (111.north) |- (121);
    \draw [->] (111.east) to (122);
    \draw [->] (111.south) |- (123);

    \node(131) at (9.5,8) [rectangle, draw=black!50, text width=\fstunequal] {\inequalitymeasure};
    \draw [->] (121) to [bend left=45] (131);
    \draw [->] (122) to (131);
    \draw [->] (123) to [bend right=45] (131);

    \node(141) at (3.00,2.4) [rectangle, draw=black!50, text width=\fstnote, label={Production function estimations}, align=left] {To illustrate, suppose that $Y_{j,i,l,y}=f\left(
    Y_{j,i,l,y-1},
    H_{j,i,l,y}\left(\left\{Z_{l,\widehat{y}}\right\}_{\widehat{y}=y-\tau}^{y}\right),
    S_{l,y}\left(\left\{Z_{l,\widehat{y}}\right\}_{\widehat{y}=y-\tau}^{y}\right),
    \left\{Z_{l,\widehat{y}}\right\}_{\widehat{y}=y-\tau}^{y},
    \epsilon_{j,i,l,y}
    \right)$, where $f(\cdot)$ is the human capital production function and $\epsilon_{j,i,l,y}$ is unobserved shock. Analysis I and II focus on understanding the relationship between environmental shocks and home and community input choices $H_{j,i,l,y}\left(\left\{Z_{l,\widehat{y}}\right\}_{\widehat{y}=y-\tau}^{y}\right)$ and $S_{l,y}\left(\left\{Z_{l,\widehat{y}}\right\}_{\widehat{y}=y-\tau}^{y}\right)$. Analysis I assumes $\tau =0$. Analysis II considers $\tau >0$, accounting for the persistent effects of past shocks. Analysis III estimates function $f(\cdot)$; potential endogeneity of inputs and unobserved shocks $\epsilon_{j,i,l,y}$ is considered, and the inclusion of $\left\{Z_{l,\widehat{y}}\right\}_{\widehat{y}=y-\tau}^{y}$ accounts for key inputs that are normally unobserved.
};
\draw [->] (131.east) to [bend left=45] (141.east);

\end{tikzpicture}
\end{figure}

\clearpage
\section{Multi-stage Choice Diagram}

Three nodes, two in a row.

\tikzset{
    % Two node styles for game trees: solid and hollow
    solid node/.style={circle,draw,inner sep=1.5,fill=black},
    hollow node/.style={circle,draw,inner sep=1.5,fill=white}
}
\begin{center}
\begin{tikzpicture}[scale=1.5,font=\footnotesize]

% A.2 Level Styles
    % Specify spacing for each level of the tree
    \tikzstyle{level 1}=[level distance=17mm,sibling distance=25mm]
    \tikzstyle{level 2}=[level distance=15mm,sibling distance=15mm]

% B Level 1 point
\node(0)[solid node,label=above:{$P1$}]{}

% C Level 2 Node (child nodes)
% C.1 First child node, left
child{
    % C.1.a Node line color
    [red]
    % C.1.b Node dot
    node(1)[solid node, label=below:{red dot}]{}
    % C.1.c now go to edge
    edge from parent
        % C.1.c Node along edge
        node[sloped, above, black, text width=3cm]{Minimum Choice showing up along this red line}
    }
% C.2 Middle Child node, do not show line, invisible
child{
    % Main node of level 2 middle node, same y-axis height as node left and right
    % Y-shift to move level 3 child node lower
    [red]  % this is the style of the line middle line
        node(2)[solid node, xshift=30, yshift=-50, label=right:{hello}]{} % this is the style of the middle point that would have matched up the middle line
    % D.1 Left level 3 (from middle level 2)
    child{
        [black]
        node[hollow node,label=below:{$(a,b)$}]{}
        edge from parent
            node[left]{$C$}
        }
    child{
        [black]
        node[hollow node,label=below:{$(c,d)$}]{}
        edge from parent
            node[right]{$D$}
        }
    edge from parent
        % draw none here prevents the middle line extending rom P1 to hello to be drawn
        [draw=none]
        %note that you need to adjust the yshift if you change the level distance
        node[left, black, xshift=-5, yshift=0, text width = 1mm]{$\alpha$\\$\beta$\\$\sigma$}
    }
child{
    [black]
    node(3)[solid node, label=right:{hi there}, green]{}
    edge from parent
        node[right, xshift=0, yshift=15, text width=2cm]{Some text for this edge}
    };
% information set
\draw[dashed, bend right]
    (1) to (3);
\draw[dashed, bend left, line width=0.5mm, blue]
    (0) to
        node[right, text width=1.25cm, xshift=5, yshift=-10]
            {\tiny{label right of blue dashed line}}
        (2);
\end{tikzpicture}
\end{center}

\section{Consumption, cash, and other spending}

\subsection{Diagram with box edge color and box fill color and parameterized layer gaps}

\def\txbwa{50}
\def\linewa{2.5}

\def\xsft{-25}
\def\lvla{0}
\def\lvlb{\lvla -1.25}
\def\lvlc{\lvlb -1.75}
\def\lvld{\lvlc -1.5}
\def\lvle{\lvld -2.25}

\def\clrsavings{orange}
\def\clrresource{gray}
\def\clrother{gray}
\def\clrcash{red}
\def\clrbank{black}

\begin{center}
\begin{tikzpicture}[thick]
    % \tikz[align=center]
    % level 1
    % \draw (0,0) rectangle (2,2) node[pos=.5, text width=2cm,  align=center] {Test sdkf ljlkdjfa lkdsjf };

    \node(101) at (0,\lvla) [rectangle,draw=\clrcash!50,fill=\clrresource!20,align=center,line width=\linewa, text width=\txbwa, xshift=\xsft, align=center] {
        Cash earn
        };
    \node(102) at (5,\lvla) [rectangle,draw=\clrcash!50,fill=\clrsavings!20, align=center, line width=\linewa, text width=\txbwa, xshift=\xsft, align=center] {
        Cash save
        };
    % level 2
    \node(201) at (2.5,\lvlb) [rectangle,draw=\clrresource!50,fill=\clrresource!20, align=center, text width=\txbwa, xshift=\xsft, align=center] {
        Resources (all cash)
        };
    % level 3
    \node(301) at (-1.5,\lvlc) [rectangle,draw=\clrother!50, fill=blue!20, align=center, text width=\txbwa, xshift=\xsft, align=center] {
        Cash consumption
        };
    \node(302) at (+2.5,\lvlc) [rectangle,draw=\clrother!50,fill=blue!20, align=center, text width=\txbwa, xshift=\xsft, align=center] {
        Non-cash cons.
        };
    \node(303) at (+6.5,\lvlc) [rectangle,draw=\clrother!50,fill=\clrsavings!20, align=center, text width=\txbwa, xshift=\xsft, align=center] {
        Savings
        };
    % level 3.5
    \node(351) at (0.075,\lvld) [rectangle,draw=\clrbank!50,fill=blue!20, line width=\linewa, text width=\txbwa, xshift=\xsft, align=center] {
        Deposit to bank
        };
    % level 4
    \node(401) at (-3.5,\lvle) [rectangle,draw=\clrcash!50,fill=blue!20, line width=\linewa, text width=\txbwa, xshift=\xsft, align=center] {
        Cash pay: Pocket-in-Pocket-out
        };
    \node(402) at (-1.1,\lvle) [rectangle,draw=\clrcash!50,fill=blue!20, line width=\linewa, text width=\txbwa, xshift=\xsft, align=center] {
        Cash pay: ATM-to-cons.
        };
    \node(403) at (1.3,\lvle) [rectangle,draw=\clrbank!50,fill=blue!20, line width=\linewa, text width=\txbwa, xshift=\xsft, align=center] {
        Bank pay: Bank-card cons.
        };
    \node(404) at (3.7,\lvle) [rectangle,draw=\clrbank!50,fill=blue!20, line width=\linewa, text width=\txbwa, xshift=\xsft, align=center] {
        Bank pay: Digital cons.
        };
    \node(405) at (6.1,\lvle) [rectangle,draw=\clrcash!50,fill=\clrsavings!20, line width=\linewa, text width=\txbwa, xshift=\xsft, align=center] {
        Cash save: Savings in pocket
        };
    \node(406) at (8.5,\lvle) [rectangle, draw=\clrbank!50, fill=\clrsavings!20, line width=\linewa, text width=\txbwa, xshift=\xsft, align=center] {
        Bank save: Savings at bank
        };

    \draw [->] (101.south) to (201.north west); 
    \draw [->] (102.south) to (201.north east);

    \draw [->] (201.south) to (301.north); 
    \draw [->] (201.south) to (302.north);
    \draw [->] (201.south) to (303.north);

    \draw [->] (301.south) to (401.north); 
    \draw [->] (301.south) to (351.north west);
    \draw [->] (302.south) to (351.north east);
    \draw [->] (351.south) to (402.north); 
    \draw [->] (351.south) to (403.north); 
    \draw [->] (351.south) to (404.north);

    \draw [->] (303.south) to (405.north);
    \draw [->] (303.south) to (406.north);

\end{tikzpicture}
\end{center}
\clearpage

\end{document}